---
title: "Larval Photosynthesis Rates Extraction with LoLinR - Bermuda 2023 ENCORE"
author: "AS Huffmyer"
output: html_document
editor_options:
  chunk_output_type: console
date: "2023-08-28"
---

Need to QC individual export plots 

This script is extracting photosynthesis rates at 410 PAR under ambient and heat conditions for P. asteroides larvae in Bermuda. The code is based on A. Huffmyer's TPC [script](https://github.com/AHuffmyer/larval_symbiont_TPC/blob/main/scripts/F1F2_sdr_extraction.Rmd).


## Setup  

Set up workspace, set options, and load required packages.
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE,message = FALSE)
```

```{r {r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
library(devtools)
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr') 
if ("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc') 
if ("respR" %in% rownames(installed.packages()) == 'FALSE') install.packages('respR') 


#load packages
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('plyr')
library('dplyr')
library('stringr')
library('Rmisc')
library('respR')
library('tidyverse')

```

## Read in files

Set the path of all respirometry files within the R project

```{r, warning=FALSE, message=FALSE}
path.p<-"data/pr_sdr/runP" #location of file
```

Bring in file names
```{r, warning=FALSE, message=FALSE}
# bring in the respiration file names
file.names<-basename(list.files(path = path.p, pattern = "csv", recursive = TRUE))
```

Bring in metadata
```{r}
run_metadata<-read.csv("data/pr_sdr/P_SDR_Run_Info.csv")
```

## Extract photosynthesis rates  

Generate photosynthesis data frames.
```{r, warning=FALSE, message=FALSE}
#generate a 6 column dataframe with specific column names
#photosynthesis
Photo.R <- data.frame(matrix(NA, ncol=7))
colnames(Photo.R) <- c("Date","Sample.ID","Run","Chamber.ID","Light", "Intercept", "umol.L.min")

Photo.Rb <- data.frame(matrix(NA, ncol=7))
colnames(Photo.Rb) <- c("Date","Sample.ID","Run","Chamber.ID","Light","Intercept",
                        "umol.L.min")

```

Load in the sample information file. It is important to have chambers in the order "A1, B1, C1, D1", rather than "A1, A2, A3, A4".  

Read in sample and run information and merge into a master metadata file. 
```{r}
# Load sample metadata (i.e., which larvae were in which runs)
Sample_Info <- read.csv(file = "data/pr_sdr/P_SDR_Sample_Info.csv")

# notes don't exist in sample..info select(!Notes)
samp <- Sample_Info$Sample_ID
run <- as.numeric(str_sub(file.names, 13, 14))
date <- str_sub(file.names,1, str_length(file.names)-11)#grab date from file name
rename <- Sample_Info$Chamber_ID


# Load run metadata (i.e., light levels and interval times for each run)
Run_Info <- read.csv(file = "data/pr_sdr/P_SDR_Run_Info.csv")
Run_Info<- Run_Info[-12,]
Light_Values <- unique(Run_Info$Light_Level)

# Join all coral and run metadata joining by date, temp and run
metadata <- full_join(Sample_Info, Run_Info) %>%
  mutate(Date = as_date(as.character(Date), format = "%Y%m%d"))

#starttime<-Run_Info$IntervalStart
#endtime<-Run_Info$IntervalStop

```

```{r}
run
Light_Values
#starttime
#endtime
date
```

Run loop to extract slopes from photosynthesis data from each light level in each file name. 
```{r, results=FALSE, warning=FALSE, message=FALSE}

for(file in 1:length(file.names))  { # for every file in list start at the first and run this following function
  
for(i in Light_Values) { #in every file, for each light value 

  Photo.Data <-read.table(file.path(path.p,file.names[file]),skip = 56, header=T, sep=",", na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Photo.Data$Temp <- Photo.Data[,17] #assigns temp column
#Photo.Data$Time.Min <- seq.int(from=0, to=((nrow(Photo.Data)*0.25)-0.25), by = 0.25) #set time in min
Photo.Data <- Photo.Data[,-c(16,21)] #remove unused columns  
  # extract start and end times for the respective plate 
  
  starttime<-Run_Info%>%
    select(Light_Level, IntervalStart)%>%
    filter(Light_Level==i)%>%
    select(IntervalStart)%>%
    as.data.frame()

   starttime<-starttime[1,1]
   
   
   endtime<-Run_Info%>%
    select(Light_Level, IntervalStop)%>%
    filter(Light_Level==i)%>%
    select(IntervalStop)%>%
    as.data.frame()

   endtime<-endtime[1,1]
   
  #filter by light interval 
  Photo.Data <- Photo.Data %>% #filters data by interval for light 
     filter(Time.Min. > starttime)%>%
     filter(Time.Min. < endtime) 

  Photo.Data.N <- Photo.Data[,3:15] #subset desired columns
  
  #add column names back in 
  Photo.Data.N<-as.data.frame(Photo.Data.N)
  
  for(j in 1:(ncol(Photo.Data.N))){
    model <- rankLocReg(
      xall=Photo.Data$Time.Min., yall=as.numeric(Photo.Data.N[, j]), 
      alpha=0.4, method="pc", verbose=TRUE) #extract slopes, percentile rank method with minimum window size of 0.4. This means that in order to fit a slope, it has to encompass at least 40% of available datapoints. 
    
  pdf(paste0("data/pr_sdr/PhotosynthesisPlots/", date[i], "_Plate",run[file], "_", rename[j], "_light", Light_Values[i], "_regression.pdf")) #Your code for generating the plot goes here

  #par("mar")
  #par(mar = c(.1, .1, .1, .1))#plot margin are to big, had to be reset
  plot(model)
  dev.off()
    
    Photo.Rb[j,1] <- as.character(date[file]) #stores the date
    Photo.Rb[j,2] <- as.character(run[file]) #stores the run number
    Photo.Rb[j,3] <- as.character(samp[j+(i-1)*ncol(Photo.Data.N)]) #stores the sample ID
    Photo.Rb[j,4] <- as.character(rename[j]) #stores the chamber ID
    Photo.Rb[j,5] <- as.character(i) #stores the chamber ID
    Photo.Rb[j,6:7] <- model$allRegs[i,c(4,5)] #inserts slope and intercept in the dataframe
    
  }
  
  Photo.R <- rbind(Photo.R, Photo.Rb) #bind final data frame
  }
  
}
```
Calculate average temperature of each run and export to a table to confirm temperatures are not different between PAR values.    

```{r}

#list files 
file.names<-basename(list.files(path = path.p, pattern = "csv", recursive = TRUE)) 

#generate matrix to populate
Temp.P <- data.frame(matrix(NA, ncol=3))
colnames(Temp.P) <- c("Date","Temp.C", "Light")

Temp.Pb <- data.frame(matrix(NA, ncol=3))
colnames(Temp.Pb) <- c("Date","Temp.C","Light")

#read in temps and generate mean values for each step of the light curve 

for(file in 1:length(file.names)){ 
  
  for(i in Light_Values) {# for every file in list start at the first and run this following function
  
    Temp.Data <-read.table(file.path(path.p,file.names[file]), skip = 56, header=T, sep=",", na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Temp.Data$Temp <- Temp.Data[,21] #assigns temp column
  Temp.Data$Time.Min <- seq.int(from=0, to=((nrow(Temp.Data)*0.25)-0.25), by = 0.25) #set time in min
  
  #filter by light interval - need to bind in dataframe to tell it which start and end time to use 
  Temp.Data <- Temp.Data %>% #filters data by interval for light 
     filter(Time.Min > starttime[i])%>%
     filter(Time.Min < endtime[i]) 
  
    Temp.Pb[j,1] <- as.character(date[file]) #stores the date
    Temp.Pb[j,3] <- mean(Temp.Data$Temp) #stores the sample ID
    Temp.Pb[j,4] <- mean(Light_Values[i]) #stores the sample ID
    
    
    Temp.P <- rbind(Temp.P, Temp.Pb) #bind final data frame
    Temp.P <- na.omit(Temp.P)
    if (ncol(Temp.P) == ncol(Temp.Pb))
  Temp.P <- rbind(Temp.P, Temp.Pb)
  
}

}
write.csv(Temp.P, paste0("output/pr_sdr/runs_temp.csv")) #save respiration rate data
plot(Photo.R$umol.L.min~as.factor(Photo.R$date)*as.factor(Photo.R$Light), side = 2, las = 2, xlab="" )

```

```{r}
