---
title: "Daily measurement analysis PAST 2023"
author: "Jill Ashey"
date: "2023-08-14"
output: html_document
---

This script reads and plots environmental data from daily measurements.  

## Set up 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

Load packages
```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
```

Read in daily measurements file.  

```{r}
daily<-read_csv("data/enviromental/daily_measurements.csv")
daily$date<-as.Date(daily$date, format="%m/%d/%y")
daily$Squarical<-as.factor(daily$Squarical)
daily$Basin <- as.factor(daily$Basin)
daily$tris.date<-as.character(daily$tris.date)

# remove values with NA for temp, salinity and pH
daily <- daily %>%
  drop_na(temp.C)
```

## Calculate total pH  

Calculate the calibration curve from the Tris calibration and calculate pH on the total scale from pH.mV.   
```{r}
pHcalib<-read_csv("data/enviromental/Tris_Calibration.csv")
pHcalib$tris.date<-as.character(pHcalib$tris.date)

pHSlope<-pHcalib %>%
  nest_by(tris.date)%>%
  mutate(fitpH = list(lm(mVTris~Ttris, data = pHcalib))) %>% # linear regression of mV and temp of the tris
  reframe(broom::tidy(fitpH)) %>% # make the output tidy
  dplyr::select(tris.date, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%# put slope and intercept in their own column
  left_join(daily, ., by="tris.date") %>% # join with the pH sample data
  mutate(mVTris = temp.C*Ttris + `(Intercept)`) %>%# calculate the mV of the tris at temperature in which the pH of the tanks were measured
  mutate(pH.total = seacarb::pH(Ex=ph.mV, Etris=mVTris, S=sal.psu, T=temp.C)) # calculate pH of the tanks using the pH seacarb function
```

pH is now calculated as Total pH in the "pH" column. Now select the desired columns to analyze further.  
```{r}
daily_calc<-pHSlope%>%
  dplyr::select(date, time, Basin, Squarical, group, temp.C, sal.psu, par, flow.mL5s, ph.nbs, pH.total)
```

## Calculate flow and par

Calculate flow to total mL per minute rather than 5 sec.  
```{r}
daily_calc<-daily_calc%>%
   mutate(flow.L.min=(flow.mL5s*12)/1000)%>%
   dplyr::select(!flow.mL5s)
```

Calculate par based on immersion correction. See Apogee manual [here](https://www.apogeeinstruments.com/underwater-par-measurements/).
```{r}
daily_calc<-daily_calc%>%
   mutate(par.adj=(par*1.15)) %>%
   dplyr::select(!par)
```

## Change to long format

Change data format to long format 
```{r}
daily_calc.long <-daily_calc %>% pivot_longer(cols=temp.C:par.adj,
  names_to = "metric",
  values_to = "value")
```

Filter by relevant dates if needed

## Plot metrics of interest  

Plot by date colored by treatment, basin and squarical.      
```{r}
## Treatment
daily_treatment<-daily_calc.long %>%
  ggplot(aes(x=date, y=value, colour=group))+
  geom_point(size=2)+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-19")), linetype="dashed", 
                #color = "black", size=0.5)+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-24")), linetype="solid", 
                #color = "black", size=0.5)+
  #scale_colour_manual(values=c("orange", "brown4"))+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  theme_bw(); daily_treatment
  
ggsave(filename="output/environmental/treatment_daily_measurements.png", plot=daily_treatment, dpi=300, width=6, height=8, units="in")

## Basin
daily_basin<-daily_calc.long %>%
  ggplot(aes(x=date, y=value, colour=Basin))+
  geom_point(size=2)+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-19")), linetype="dashed", 
                #color = "black", size=0.5)+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-24")), linetype="solid", 
                #color = "black", size=0.5)+
  #scale_colour_manual(values=c("orange", "brown4"))+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  theme_bw(); daily_basin
  
ggsave(filename="output/environmental/basin_daily_measurements.png", plot=daily_basin, dpi=300, width=6, height=8, units="in")

## Squarical
daily_squarical<-daily_calc.long %>%
  ggplot(aes(x=date, y=value, colour=Squarical))+
  geom_point(size=2)+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-19")), linetype="dashed", 
                #color = "black", size=0.5)+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-24")), linetype="solid", 
                #color = "black", size=0.5)+
  #scale_colour_manual(values=c("orange", "brown4"))+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  theme_bw(); daily_squarical
  
ggsave(filename="output/environmental/squarical_daily_measurements.png", plot=daily_squarical, dpi=300, width=6, height=8, units="in")
```

Plot by treatment summarizing tanks within each group    
```{r}
daily_treatment<-daily_calc.long %>%
  ggplot(aes(x=date, y=value, colour=group, group=interaction(Basin, date)))+
  geom_point()+
  geom_boxplot()+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-19")), linetype="dashed", 
                #color = "black", size=0.5)+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-24")), linetype="solid", 
                #color = "black", size=0.5)+
  #scale_colour_manual(values=c("orange", "brown4"))+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  theme_bw(); daily_treatment

ggsave(filename="output/environmental/boxplot_daily_measurements.png", plot=daily_treatment, dpi=300, width=6, height=8, units="in")
```








