---
title: "Daily measurement analysis PAST 2023"
author: "Jill Ashey"
date: "2023-08-14"
output: html_document
---

This script reads and plots environmental data from daily measurements.  

## Set up 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

Load packages
```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
```

Read in daily measurements file.  

```{r}
daily<-read_csv("data/enviromental/daily_measurements.csv")
daily$date<-as.Date(daily$date, format="%m/%d/%y")
daily$tris.date<-as.character(daily$tris.date)

# remove values with NA for temp, salinity and pH
daily <- daily %>%
  drop_na(temp.C)
```

## Calculate total pH  

Calculate the calibration curve from the Tris calibration and calculate pH on the total scale from pH.mV.   
```{r}
pHcalib<-read_csv("data/enviromental/Tris_Calibration.csv")
pHcalib$tris.date<-as.character(pHcalib$tris.date)

pHSlope<-pHcalib %>%
  nest_by(tris.date)%>%
  mutate(fitpH = list(lm(mVTris~Ttris, data = pHcalib))) %>% # linear regression of mV and temp of the tris
  reframe(broom::tidy(fitpH)) %>% # make the output tidy
  dplyr::select(tris.date, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%# put slope and intercept in their own column
  left_join(daily, ., by="tris.date") %>% # join with the pH sample data
  mutate(mVTris = temp.C*Ttris + `(Intercept)`) %>%# calculate the mV of the tris at temperature in which the pH of the tanks were measured
  mutate(pH.total = seacarb::pH(Ex=ph.mV, Etris=mVTris, S=sal.psu, T=temp.C)) # calculate pH of the tanks using the pH seacarb function
```

pH is now calculated as Total pH in the "pH" column. Now select the desired columns to analyze further.  
```{r}
daily_calc<-pHSlope%>%
  dplyr::select(date, time, Type, group, temp.C, sal.psu, par, flow.mL5s, ph.nbs, pH.total)
```

## Calculate flow and par

Calculate flow to total mL per minute rather than 5 sec.  
```{r}
daily_calc<-daily_calc%>%
   mutate(flow.L.min=(flow.mL5s*12)/1000)%>%
   dplyr::select(!flow.mL5s)
```

Calculate par based on immersion correction. See Apogee manual [here](https://www.apogeeinstruments.com/underwater-par-measurements/).
```{r}
daily_calc<-daily_calc%>%
   mutate(par.adj=(par*1.15)) %>%
   dplyr::select(!par)
```

## Change to long format

Change data format to long format 
```{r}
daily_calc.long <-daily_calc %>% pivot_longer(cols=temp.C:par.adj,
  names_to = "metric",
  values_to = "value")
```

Filter by relevant dates if needed

## Plot metrics of interest  

Plot by date colored by treatment, basin and squarical.      
```{r}
## Treatment
daily_treatment<-daily_calc.long %>%
  ggplot(aes(x=date, y=value, colour=group))+
  geom_point(size=2)+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-19")), linetype="dashed", 
                #color = "black", size=0.5)+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-24")), linetype="solid", 
                #color = "black", size=0.5)+
  #scale_colour_manual(values=c("orange", "brown4"))+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  theme_bw(); daily_treatment
  
ggsave(filename="output/environmental/treatment_daily_measurements.png", plot=daily_treatment, dpi=300, width=6, height=8, units="in")

## Tank type
daily_tank<-daily_calc.long %>%
  ggplot(aes(x=date, y=value, colour=Type))+
  geom_point(size=2)+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-19")), linetype="dashed", 
                #color = "black", size=0.5)+
  #geom_vline(xintercept = as.numeric(ymd("2022-10-24")), linetype="solid", 
                #color = "black", size=0.5)+
  #scale_colour_manual(values=c("orange", "brown4"))+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  theme_bw(); daily_tank
  
ggsave(filename="output/environmental/tank_daily_measurements.png", plot=daily_tank, dpi=300, width=6, height=8, units="in")
```

Plot means by basin and squarical.      
```{r}
# Plot by basin
basin_daily_treatment_summary<-daily_calc.long %>%
  #filter(is.na(Squarical)) %>%
  filter(Type %in% c("Basin 1", "Basin 2", "Basin 3", "Basin 4", "Basin 5", "Basin 6"))%>%
  
  ggplot(aes(x=Type, y=value, colour=group))+
  geom_point()+
  geom_boxplot()+
  scale_colour_manual(name="", values=c("blue", "red"))+
  xlab("Date")+
  ylab("Value")+
  facet_wrap(~ metric, scales = "free")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  # Rotate text by 45 degrees
  theme_classic(); basin_daily_treatment_summary

ggsave(filename="output/environmental/basin_summary_daily_measurements.png", plot=basin_daily_treatment_summary, dpi=300, width=9, height=6, units="in")

# Plot by squarical
squarical_daily_treatment_summary<-daily_calc.long %>%
  #filter(is.na(Squarical)) %>%
  filter(Type %in% c("Squarical 1", "Squarical 2", "Squarical 3", "Squarical 4", "Squarical 5", "Squarical 6", "Squarical 7", "Squarical 8", "Squarical 9", "Squarical 10", "Squarical 11", "Squarical 12", "Squarical 13", "Squarical 14", "Squarical 15", "Squarical 16"))%>%
  
  ggplot(aes(x=Type, y=value, colour=group))+
  geom_point()+
  geom_boxplot()+
  scale_colour_manual(name="", values=c("blue", "red"))+
  xlab("Date")+
  ylab("Value")+
  facet_wrap(~ metric, scales = "free")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  # Rotate text by 45 degrees
  theme_classic(); squarical_daily_treatment_summary

ggsave(filename="output/environmental/squarical_summary_daily_measurements.png", plot=squarical_daily_treatment_summary, dpi=300, width=9, height=6, units="in")
```

## Test for differences   

Use the daily_calc df created earlier.

Test for differences between treatments.
```{r}
daily_calc%>%
  aov(flow.L.min~group, data=.)%>%
  summary()
 
daily_calc%>%
  aov(par.adj~group, data=.)%>%
  summary()

daily_calc%>%
  aov(pH.total~group, data=.)%>%
  summary()

daily_calc%>%
  aov(ph.nbs~group, data=.)%>%
  summary()

daily_calc%>%
  aov(sal.psu~group, data=.)%>%
  summary()
# significantly different by treatment

daily_calc%>%
  aov(temp.C~group, data=.)%>%
  summary()
# significantly different by treatment
```

Test for differences between tank type and treatment 
```{r}
daily_calc%>%
  #filter(Type %in% c("Basin 1", "Basin 2", "Basin 3", "Basin 4", "Basin 5", "Basin 6"))%>%
  aov(flow.L.min~Type*group, data=.)
 
daily_calc%>%
  aov(par.adj~Type*group, data=.)%>%
  summary()

daily_calc%>%
  aov(pH.total~Type*group, data=.)%>%
  summary()
# significantly different by tank type

daily_calc%>%
  aov(ph.nbs~Type*group, data=.)%>%
  summary()
# significantly different by tank type

daily_calc%>%
  aov(sal.psu~Type*group, data=.)%>%
  summary()

daily_calc%>%
  aov(temp.C~Type*group, data=.)%>%
  summary()
# significantly different by tank type and treatment
```

Remove Basins from tank Type and test differences by squarical  
```{r}
squarical_daily <- daily_calc %>%
  dplyr::filter(!str_detect(Type, "Basin"))

# Run stats 
squarical_daily%>%
  aov(flow.L.min~Type, data=.) %>%
  summary()
 
squarical_daily%>%
  aov(par.adj~Type, data=.) %>%
  summary()

squarical_daily%>%
  aov(pH.total~Type, data=.) %>%
  summary()

squarical_daily%>%
  aov(ph.nbs~Type, data=.) %>%
  summary() 

squarical_daily%>%
  aov(sal.psu~Type, data=.) %>%
  summary()

squarical_daily%>%
  aov(temp.C~Type, data=.) %>%
  summary()
```

## Summarize daily measurements  

Calculate descriptive statistics   
```{r}
summary<-daily_calc%>%
  group_by(Type, group)%>%
  #select(!tank)%>%
  select(!date)%>%
  select(!time)%>%
  summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE)); summary

write_csv(summary, "output/environmental/daily_measurements_summary.csv")
```




