---
output: html_document
editor_options: 
  chunk_output_type: console
---
--
title: "Larval Photosynthesis Rates Extraction with LoLinR - Bermuda 2023 - PI Curves"
author: "Jill Ashey"
date: "2023-07-13"
output: html_document
---

This script is extracting photosynthesis rates under different light levels for P. asteroides larvae in Bermuda. The code is based on A. Huffmyer's PI [script](https://github.com/AHuffmyer/larval_symbiont_TPC/blob/main/scripts/larval_pi_curve_sdr_extraction.Rmd). Important note - this data was collected in 4 mL vials, not the 24-well plates.

## Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
library(devtools)
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr') 
if ("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc') 
if ("respR" %in% rownames(installed.packages()) == 'FALSE') install.packages('respR') 


#load packages
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('plyr')
library('dplyr')
library('stringr')
library('Rmisc')
library('respR')
library("tidyverse")
```

## Read in files  

Set the path of all respirometry files within the R project.   

```{r, warning=FALSE, message=FALSE}
path.p<-"data/larval_pi_curves/runs" #location of files
```

Bring in the file names.
```{r, warning=FALSE, message=FALSE}
# bring in the respiration file names
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) 
# remove test file once we have more pi curve data 
```

Bring in the run metadata. 
```{r}
run_metadata<-read.csv("data/larval_pi_curves/larval_PIcurve_SDR_Run_Info.csv")
```

## Extract photosynthesis rates  

Generate photosynthesis data frames.

```{r, warning=FALSE, message=FALSE}
#generate a 7 column dataframe with specific column names
#photosynthesis
Photo.R <- data.frame(matrix(NA, ncol=7))
colnames(Photo.R) <- c("Date", "Plate","Sample.ID","Chamber.ID","Light", "Intercept", "umol.L.min")

Photo.Rb <- data.frame(matrix(NA, ncol=7))
colnames(Photo.Rb) <- c("Date", "Plate","Sample.ID","Chamber.ID","Light", "Intercept", "umol.L.min")
```

Load in the sample information file. It is important to have chambers in the order "A1, B1, C1, D1", rather than "A1, A2, A3, A4".  

Read in sample and run information and merge into a master metadata file. 
```{r}
# Load PI curve sample metadata (i.e., which larvae were in which runs)
Sample.Info <- read_csv(file = "data/larval_pi_curves/larval_PIcurve_SDR_Sample_Info.csv")%>%
  select(!TimeStart)%>%
  select(!TimeEnd)%>%
  select(!Notes)

rename <- Sample.Info$Chamber.ID
samp <- Sample.Info$Sample.ID

plate <- as.numeric(str_sub(file.names, 28, 29))
date <- str_sub(file.names, 1, str_length(file.names)-28) #grab date from file name

# Load PI curve run metadata (i.e., light levels and interval times for each run)
Run.Info <- read_csv(file = "data/larval_pi_curves/larval_PIcurve_SDR_Run_Info.csv")

Light_Values <- Run.Info$Light_Level

# Join all coral and run metadata
metadata <- full_join(Sample.Info, Run.Info) %>%
  mutate(Date = as_date(as.character(Date), format = "%Y%m%d"))

starttime<-Run.Info$IntervalStart
endtime<-Run.Info$IntervalStop
```

Script is having an issue extracting the plate number from the file name. 

Remove "Type" columns that have NA in them 
```{r}
metadata <- metadata %>%
  drop_na(Type)
```


Run loop to extract slopes from photosynthesis data from each light level in each file name/plate. 

```{r, results=FALSE, warning=FALSE, message=FALSE}

for(file in 1:length(file.names)) { # for every file in list start at the first and run this following function
  
  for (i in Light_Values) { #in every file, for each light value 
    
  Photo.Data <-read.table(file.path(path.p,file.names[file]), skip = 56, header=T, sep=",", na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Photo.Data$Temp <- Photo.Data[,31] #assigns temp column
  Photo.Data$Time.Min. <- seq.int(from=0, to=((nrow(Photo.Data)*0.25)-0.25), by = 0.25) #set time in min
  
  # extract start and end times for the respective plate 
  
  starttime<-Run.Info%>%
    select(Plate, Light_Level, IntervalStart)%>%
    filter(Plate==plate[file])%>%
    filter(Light_Level==i)%>%
    select(IntervalStart)%>%
    as.data.frame()

   starttime<-starttime[1,1]
   
   endtime<-Run.Info%>%
    select(Plate, Light_Level, IntervalStop)%>%
    filter(Plate==plate[file])%>%
    filter(Light_Level==i)%>%
    select(IntervalStop)%>%
    as.data.frame()

   endtime<-endtime[1,1]
   
  #filter by light interval - need to bind in dataframe to tell it which start and end time to use 
  Photo.Data <- Photo.Data %>% #filters data by interval for light 
     filter(Time.Min. > starttime)%>%
     filter(Time.Min. < endtime) 
  
  Photo.Data.N <- Photo.Data[,3:26] #subset desired columns
  
  #add column names back in 
  Photo.Data.N<-as.data.frame(Photo.Data.N)
  
  for(j in 1:(ncol(Photo.Data.N))){
    model <- rankLocReg(
      xall=Photo.Data$Time.Min., yall=as.numeric(Photo.Data.N[, j]), 
      alpha=0.4, method="pc", verbose=TRUE) #extract slopes, percentile rank method with minimum window size of 0.4. This means that in order to fit a slope, it has to encompass at least 40% of available datapoints. 
    
    pdf(paste0("output/larval_pi_curves/PhotosynthesisPlots/",date[file], "_Plate",plate[file],"_",rename[j],"_light", Light_Values[i],"_regression_trunc.pdf")) #generate output file names
    plot(model)
    dev.off()
    
        Photo.Rb[j,1] <- as.character(date[file]) #stores the date
    Photo.Rb[j,2] <- as.character(plate[file]) #stores the run number
    Photo.Rb[j,3] <- as.character(samp[j+(i-1)*ncol(Photo.Data.N)]) #stores the sample ID
    Photo.Rb[j,4] <- as.character(rename[j]) #stores the chamber ID
    Photo.Rb[j,5] <- as.character(i) #stores the chamber ID
    Photo.Rb[j,6:7] <- model$allRegs[i,c(4,5)] #inserts slope and intercept in the dataframe
    
  }
  
  Photo.R <- rbind(Photo.R, Photo.Rb) #bind final data frame
  }
  
}
```

There is some issue with the start/end times and the Time.Min. I think the issue is somewhere between lines 134-157, but I'm not totally sure. Every time I try to run the chunk above, it gives me this error: "Error in combn(lenX, 2) : n < m", which makes me think that the issue is with the time filtering step. Again, not totally sure. Might be best to talk to AH about this. 

^^ AH: The reason for this problem is that there is no data available in the data frame. If you look at the data frame its running all wells have "no sensor" in them. So there is no data for it to calculate.


   


