---
title: "Normalizing Rates"
author: "HM Putnam"
date: "2023-11-26"
output: html_document
---

## Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
library(devtools)
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr') 
if ("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc') 
if ("respR" %in% rownames(installed.packages()) == 'FALSE') install.packages('respR') 
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("ggpubr" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggpubr') 

#load packages
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('plyr')
library('dplyr')
library('stringr')
library('Rmisc')
library('respR')
library('tidyverse')
library('emmeans')
library('ggpubr')

```

Load Raw extracted Oxygen Flux Rates
```{r, warning=FALSE, message=FALSE}
Photo.Rates <- read.csv(file = "output/photo_sdr_rates_raw.csv") #read file back in so slopes don't have to be generated every time

#format  column
Photo.Rates<-Photo.Rates %>%
  mutate(Plate = str_sub(Plate, 1, -1))
Photo.Rates$Plate<-as.integer(Photo.Rates$Plate) #format as # rather than run #, set as integer
```

## Standardize and normalize  

Merge rates files with sample info for normalization, plotting, and stats.  

```{r, warning=FALSE, message=FALSE}
Sample.Info <- read_csv(file = "data/pr_sdr/PR_SDR_Sample_Info.csv")

#Sample.Info <- Sample.Info %>%
#  filter(Date == 20231216)

#make unique ID column
Sample.Info$unique <- paste0(Sample.Info$Date, Sample.Info$Plate, Sample.Info$Chamber_ID)

Photo.Rates$unique <- paste0(Photo.Rates$Date, Photo.Rates$Plate, Photo.Rates$Chamber_ID)

#oxygen<-left_join(Sample.Info, Photo.Rates, by="unique") #add photosynthesis data
oxygen<- left_join(Sample.Info, Photo.Rates, by = "unique", suffix = c("", ".y")) 

colnames(oxygen)[colnames(oxygen) == 'Intercept'] <- 'Photo.Intercept' #rename to specify R
colnames(oxygen)[colnames(oxygen) == 'umol.L.min'] <- 'P.umol.L.min' #rename to specify R

oxygen<- oxygen %>%
  filter(Type=="Blank"|Type=="Sample")

```

Check for Outliers in the raw data
```{r}
oxygen %>% 
  filter(Type=="Blank") %>%
  ggplot(aes(Phenotype, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

oxygen %>% 
  filter(Type=="Sample") %>%
  ggplot(aes(Phenotype, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

oxygen %>% 
  ggplot(aes(Type, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Date, Light, Phenotype), ncol=4)+
  theme_bw()


```

Remove samples that have inaccurate slope extraction for respiration rates (determined by PDF output files and outlier plots and data filled in for "No Sensor" as a test of the code function )  
```{r}
oxygen<-oxygen%>%
  filter(!(unique=="202311174D5")) %>%
  filter(!(unique=="202311175D5")) %>%
  filter(!(unique=="202311195D5")) %>%
  filter(!(unique=="202311176D5")) %>%
  filter(!(unique=="202311196D5")) %>%
  filter(!(unique=="202312137D5"))

oxygen %>% 
  filter(Type=="Blank") %>%
  ggplot(aes(Date, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

oxygen %>% 
  filter(Type=="Sample") %>%
  ggplot(aes(Date, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

oxygen %>% 
  ggplot(aes(Type, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Date, Light, Phenotype), ncol=4)+
  theme_bw()

```

Account for volume to obtain umol per minute.   
```{r, results=TRUE, warning=FALSE, message=FALSE}
#Account for chamber volume to convert from umol L-1 m-1 to umol m-1. This removes per Liter
oxygen$P.umol.min <- oxygen$P.umol.L.min * oxygen$Volume #calculate

```

Check the data from the test of the Number of Larvae needed to get a signal
```{r, results=TRUE, warning=FALSE, message=FALSE}

#filter test data
	
test.number.data <- oxygen %>%
  filter(Date==20231113)

#Subtract blank values. Average blank calculated for each run. Display mean blank 
test_blank_data <- subset(test.number.data, Type == "Blank") #subset to blank data only

test_blank_data %>% 
  ggplot(aes(Light, P.umol.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme_bw()

#remove high outlier blanks 
test_blank_data <- test_blank_data %>%
  filter(!P.umol.min>1.0e-04)

#remove low outlier blanks 
test_blank_data <- test_blank_data %>%
  filter(!P.umol.min < (-1.0e-04))

test_blank_data %>% 
  ggplot(aes(Light, P.umol.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme_bw()

#display mean blank values
mean.test.blank.data <- test_blank_data %>% group_by(Light) %>%
summarise(mean = mean(P.umol.min, na.rm=TRUE)) #mean P phase blanks

mean.test.blank.data %>% 
  ggplot(aes(Light, mean))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme_bw()


test.photo.blnk <- aggregate(P.umol.min ~ Light, data=test_blank_data, mean) #calculate average blank during light for each run
colnames(test.photo.blnk)[colnames(test.photo.blnk) == 'P.umol.min'] <- 'P.Blank.umol.min' #rename to specify blank for R

test.number.data.oxygen <- full_join(test.number.data, test.photo.blnk) #add R blanks to master

test.number.data.oxygen$P.umol.min.corr<-test.number.data.oxygen$P.umol.min-test.number.data.oxygen$P.Blank.umol.min #subtract R blanks

#normalize to larval number
#respiration
test.number.data.oxygen$P.umol.org.min <- test.number.data.oxygen$P.umol.min.corr/test.number.data.oxygen$Org.Number #calculate oxygen per organism 
test.number.data.oxygen$P.nmol.org.min <- test.number.data.oxygen$P.umol.org.min*1000 #calculate nanomoles 

test.number.data.oxygen %>%
  filter(Type=="Sample")%>%
  ggplot(aes(Org.Number, P.umol.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()


#60 larvae generates measureable signals for dark respiration (1) and Photosynthesis (2)

```

Analyze the data by Cohorts without the data for testing by number of larvae 
```{r}
#filter to cohort data
oxygen <- oxygen %>%
  filter(Date==20231216) #%>%
  #filter(!Date==20231113) %>%
  #filter(!Temperature==36)

blank_data <- subset(oxygen, Type == "Blank") #subset to blank data only

blank_data %>% 
  ggplot(aes(Cohort, P.umol.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

#remove high outlier blanks 
blank_data <- blank_data %>%
  filter(!P.umol.min>1.0e-04)

#remove low outlier blanks 
blank_data <- blank_data %>%
  filter(!P.umol.min < (-1.0e-04))

blank_data %>% 
  ggplot(aes(Cohort, P.umol.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

#display mean blank values
mean.blank.data <- blank_data %>% group_by(Plate, Light) %>%
summarise(mean = mean(P.umol.min, na.rm=TRUE)) #mean P phase blanks

mean.blank.data %>% 
  ggplot(aes(Plate, mean))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()


photo.blnk <- aggregate(P.umol.min ~ Date*Plate*Light, data=blank_data, mean) #calculate average blank during light for each run
colnames(photo.blnk)[colnames(photo.blnk) == 'P.umol.min'] <- 'P.Blank.umol.min' #rename to specify blank for R

oxygen <- full_join(oxygen, photo.blnk) #add R blanks to master
```

Subtract blank values to generate a "corrected" value for umol O2 min-1.  

```{r, warning=FALSE, message=FALSE}
oxygen$P.umol.min.corr<-oxygen$P.umol.min-oxygen$P.Blank.umol.min #subtract R blanks
```

Normalize to biologically relevant measure. Here, normalize to number of larvae. This can be substituted or changed for larval size/volume as well.  

```{r, warning=FALSE, message=FALSE}
oxygen.bio <- oxygen %>% filter(Type == 60) #isolate only biological samples and 
#respiration
oxygen.bio$P.umol.org.min <- oxygen.bio$P.umol.min.corr/oxygen.bio$Org.Number #calculate oxygen per organism 
oxygen.bio$P.nmol.org.min <- oxygen.bio$P.umol.org.min*1000 #calculate nanomoles 

```

Plot values.  
```{r}
options(scipen = 999)
oxygen.bio %>% 
  ggplot(aes(Cohort, P.umol.min.corr))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

oxygen.bio$Cohort_Phenotype <- paste0(oxygen.bio$Cohort, "  ", oxygen.bio$Phenotype)

options(scipen = 999)
oxygen.bio %>% 
  ggplot(aes(Cohort_Phenotype, P.umol.min.corr))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

```


2-way ANOVA on Cohorts of green fluorescing larvae
```{r}

#filter to the green larvae only
oxygen.bio.green <- oxygen.bio %>%
  filter(!Phenotype == "Red")

oxygen.bio.green %>% 
  ggplot(aes(Cohort, P.umol.min.corr))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

#subset into photo data and resp data
Photo.bio.green <- oxygen.bio.green %>%
  filter(Light == 1) 

Photo.bio.green <- Photo.bio.green %>%
  filter(!unique == "202311196A1")


mod1 <- lm(P.nmol.org.min ~ as.factor(Cohort), data=Photo.bio.green )
summary(mod1)
hist(mod1$residuals)
plot(mod1$fitted.values, mod1$residuals)

posthoc1 <- emmeans(mod1, pairwise ~ Cohort)
posthoc1

#Respiration Analysis
Resp.bio.green <- oxygen.bio.green %>%
  filter(Light == 2)


Resp.bio.green <- Resp.bio.green %>%
  filter(!unique == "202311196A1")

#remove positive values for respiration
#Resp.bio.green <- Resp.bio.green %>%
#  filter(P.nmol.org.min < 0 )

mod2 <- lm(P.nmol.org.min ~as.factor(Cohort), data=Resp.bio.green )
summary(mod2)
hist(mod2$residuals)
plot(mod2$fitted.values, mod2$residuals)
posthoc2 <- emmeans(mod2, pairwise ~ Cohort)
posthoc2

#Pgross = Pnet - -Rdark

oxygen.bio.green$group <- paste0(oxygen.bio.green$Date, "_",oxygen.bio.green$Chamber.ID)

Pgross.calcs <- as.data.frame(cbind(Resp.bio.green$unique, Resp.bio.green$Cohort, Resp.bio.green$P.nmol.org.min,Photo.bio.green$unique, Photo.bio.green$P.nmol.org.min))

colnames(Pgross.calcs) <- c("Resp.Ids", "Cohort","Resp.P.nmol.org.min","Photo.Ids", "Photo.P.nmol.org.min")

Pgross.calcs$check <- ifelse(Pgross.calcs$Resp.Ids == Pgross.calcs$Photo.Ids, 'True', 
                      ifelse(data$Resp.Ids != data$Photo.Ids, "False"))

Pgross.calcs$Resp.P.nmol.org.min <- as.numeric(as.character(Pgross.calcs$Resp.P.nmol.org.min))
Pgross.calcs$Photo.P.nmol.org.min <- as.numeric(as.character(Pgross.calcs$Photo.P.nmol.org.min))


Pgross.calcs$Pgross <-Pgross.calcs$Photo.P.nmol.org.min-Pgross.calcs$Resp.P.nmol.org.min
Pgross.calcs$PtoR <-Pgross.calcs$Pgross/(abs(Pgross.calcs$Resp.P.nmol.org.min))

plot.oxygen.bio.green <- rbind(Photo.bio.green,Resp.bio.green)
# supp.labs <- c("Photosynthesis (Pnet)", "Respiration (LEDR)")
# names(supp.labs) <- c("1", "2")

larval.photo <- Photo.bio.green %>% ggplot(., aes(x = Cohort, y = P.nmol.org.min)) +
  geom_point(data = Photo.bio.green, aes(Cohort, P.nmol.org.min), color="grey") +
  stat_summary(fun=mean, shape=1, size=0.5, color="black") +
  stat_summary(fun.data=mean_se, color = 'black', geom = "errorbar", width=0)+
  annotate("text", x = c(20231112.2,20231113.2,20231115.2), y = c(0.00215, 0.0011, -0.0005), label = c("A", "A", "B"))+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  ylim(-0.003,0.005)+
  ylab("Photosynthesis (Pnet, nmol larva-1 min-1)")+
  theme_bw()+
  theme(axis.title.x=element_blank())
larval.photo

larval.resp <- Resp.bio.green %>% ggplot(., aes(x = Cohort, y = P.nmol.org.min)) +
  geom_point(data = Resp.bio.green, aes(Cohort, P.nmol.org.min), color="grey") +
  stat_summary(fun=mean, shape=1, size=0.5, color="black") +
  stat_summary(fun.data=mean_se, color = 'black', geom = "errorbar", width=0)+
  annotate("text", x = c(20231112.2,20231113.2,20231115.2), y = c(-0.0023, -0.00095, -0.0019), label = c("A", "B", "AB"))+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  ylim(-0.003,0.005)+
  ylab("Respiration (LEDR, nmol larva-1 min-1)")+
  theme_bw()+
  theme(axis.title.x=element_blank())
larval.resp

mod3 <- lm(Pgross ~as.factor(Cohort), data=Pgross.calcs )
summary(mod3)
hist(mod3$residuals)
plot(mod3$fitted.values, mod3$residuals)
posthoc3 <- emmeans(mod3, pairwise ~ Cohort)
posthoc3

larval.Pgross <- Pgross.calcs %>% ggplot(., aes(x = as.numeric(Cohort), y = Pgross)) +
  geom_point(data = Pgross.calcs, aes(as.numeric(Cohort), Pgross), color="grey") +
  stat_summary(fun=mean, shape=1, size=0.5, color="black") +
  stat_summary(fun.data=mean_se, color = 'black', geom = "errorbar", width=0)+
  annotate("text", x = c(20231112.2,20231113.2,20231115.2), y = c(0.0035, 0.0018, 0.0017), label = c("A", "B", "B"))+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  ylim(-0.003,0.005)+
  ylab("Photosynthesis (Pgross, nmol larva-1 min-1)")+
  theme_bw()+
  theme(axis.title.x=element_blank())
larval.Pgross

mod4 <- lm(PtoR ~as.factor(Cohort), data=Pgross.calcs )
summary(mod4)
hist(mod4$residuals)
plot(mod4$fitted.values, mod4$residuals)
posthoc4 <- emmeans(mod4, pairwise ~ Cohort)
posthoc4

larval.PtoR <- Pgross.calcs %>% ggplot(., aes(x = as.numeric(Cohort), y = PtoR)) +
  geom_point(data = Pgross.calcs, aes(as.numeric(Cohort), PtoR), color="grey") +
  stat_summary(fun=mean, shape=1, size=0.5, color="black") +
  stat_summary(fun.data=mean_se, color = 'black', geom = "errorbar", width=0)+
  annotate("text", x = c(20231112.2,20231113.2,20231115.2), y = c(2.1, 2.3, 1), label = c("A", "A", "B"))+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  ylim(-6,6)+
  ylab("Pgross : LEDR")+
  theme_bw()+
  theme(axis.title.x=element_blank())
larval.PtoR


larval.respo <- ggarrange(larval.photo, larval.resp, 
                          larval.Pgross, larval.PtoR, ncol = 4, nrow = 1)

ggsave("output/Nov_NewMoon_Larval_Metabolism.pdf", larval.respo, width = 12, height = 4)
ggsave("output/Nov_NewMoon_Larval_Metabolism.jpg", larval.respo, width = 12, height = 4)

```

Save as .csv file.  
```{r, warning=FALSE, message=FALSE}
write.csv(oxygen.bio.green, paste0("output/cohort_pr_calculated_normalized_photo_rates.csv")) #save final file
```


Comparison of Green and Red Phenotypes
```{r, warning=FALSE, message=FALSE}
oxygen.bio.phenotypes <- oxygen.bio 

cohort.pheno <- oxygen.bio.phenotypes %>% 
  ggplot(aes(as.factor(Light), P.nmol.org.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  scale_x_discrete(labels=c("1" = "Pnet", "2" = "LEDR"),name ="Assay")+
  theme(axis.text.x = element_blank())+
  facet_wrap(vars(Cohort_Phenotype), ncol=4)+
  ylab("Oxygen Flux (nmol larva-1 min-1)")+
  theme_bw()
cohort.pheno 
ggsave("output/Nov_NewMoon_Phenotype_Cohort_Metabolism.pdf", cohort.pheno , width = 6, height = 4)
ggsave("output/Nov_NewMoon_Phenotype_Cohort_Metabolism.jpg", cohort.pheno , width = 6, height = 4)


```

Comparison of development at 28 and 31 
```{r, warning=FALSE, message=FALSE}
oxygen.bio.devo <- oxygen.bio %>%
  filter(Date == 20231213)

devo.temp <- oxygen.bio.devo %>% 
  ggplot(aes(as.factor(Temperature), P.nmol.org.min, group=Phenotype, color=Phenotype))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  scale_x_discrete(labels=c("1" = "Pnet", "2" = "LEDR"),name ="Assay")+
  theme(axis.text.x = element_blank())+
  facet_wrap(vars(Light), ncol=4)+
  ylab("Oxygen Flux (nmol larva-1 min-1)")+
  theme_bw()
cohort.pheno 
ggsave("output/Nov_NewMoon_Phenotype_Cohort_Metabolism.pdf", cohort.pheno , width = 6, height = 4)
ggsave("output/Nov_NewMoon_Phenotype_Cohort_Metabolism.jpg", cohort.pheno , width = 6, height = 4)


oxygen.bio.devo$group <- paste0(oxygen.bio.devo$Temperature, oxygen.bio.devo$Phenotype)

ggplot(aes(y = P.nmol.org.min, x = group, fill = Phenotype), data = oxygen.bio.devo) + geom_boxplot() + facet_wrap(vars(Light), ncol=4)


```
```{r}
ggplot(oxygen.bio.devo, aes(x = Temperature, y = P.nmol.org.min)) +
  stat_summary(aes(group = Phenotype), fun.y = mean, geom = "path",  position = position_dodge(width = 0.3)) +
  stat_summary(aes(color = Phenotype), fun.data = mean_cl_boot, geom = "errorbar", width = 0.1,  position = position_dodge(width = 0.3)) +
  stat_summary(aes(color = Phenotype), fun.y = mean, geom = "point", size = 4,  position = position_dodge(width = 0.3)) +
  geom_point(aes(color = Phenotype), size = 2, position = position_dodge(width = 0.3)) +
  facet_wrap(vars(Light), ncol=4) +
  scale_color_manual(values = c("blue", "red"))+
  theme_bw()


oxygen.bio.devo.light <- oxygen.bio.devo %>%
  filter(Light == 1)

model.devo.light <- lm(P.nmol.org.min~ Temperature*Phenotype, data=oxygen.bio.devo.light)
summary(model.devo.light)

oxygen.bio.devo.dark <- oxygen.bio.devo %>%
  filter(Light == 2)

model.devo.dark <- lm(P.nmol.org.min~ Temperature*Phenotype, data=oxygen.bio.devo.dark)
summary(model.devo.dark)

```