---
title: "Larval Photosynthesis Rates Extraction with LoLinR - Moorea Pocillopora 2023 - PR Measurements"
author: "AS Huffmyer Hollie Putnam"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

Need to QC individual export plots 

## Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
library(devtools)
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr') 
if ("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc') 
if ("respR" %in% rownames(installed.packages()) == 'FALSE') install.packages('respR') 


#load packages
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('plyr')
library('dplyr')
library('stringr')
library('Rmisc')
library('respR')
library('tidyverse')

```

## Read in files  

Set the path of all respirometry files within the R project.   

```{r, warning=FALSE, message=FALSE}
path.p<-"data/pr_sdr/runR" #location of files
```

Bring in the file names.
```{r, warning=FALSE, message=FALSE}
# bring in the respiration file names
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) 
```

Bring in the run metadata. 
```{r}
run_metadata<-read.csv("data/pr_sdr/R_SDR_Run_Info.csv")
```

## Extract respiration rates  

Generate respiration data frames.

```{r, warning=FALSE, message=FALSE}
#generate a 7 column dataframe with specific column names
#photosynthesis
Resp.R <- data.frame(matrix(NA, ncol=6))
colnames(Resp.R) <- c("Date", "Plate","Chamber.ID","Light", "Intercept", "umol.L.min")

Resp.Rb <- data.frame(matrix(NA, ncol=6))
colnames(Resp.Rb) <- c("Date", "Plate","Chamber.ID","Light", "Intercept", "umol.L.min")

```

Load in the sample information file. It is important to have chambers in the order "A1, B1, C1, D1", rather than "A1, A2, A3, A4".  

Read in sample and run information and merge into a master metadata file. 
```{r}
# Load sample metadata (i.e., which larvae were in which runs)
Sample.Info <- read_csv(file = "data/pr_sdr/PR_SDR_Sample_Info.csv") #%>%
  #select(!Notes)

rename <- Sample.Info$Chamber_ID
samp <- Sample.Info$Sample_ID

plate <- as.numeric(str_sub(file.names, 26, 27))
date <- str_sub(file.names, 4, str_length(file.names)-26) #grab date from file name

# Load PI curve run metadata (i.e., light levels and interval times for each run)
Run.Info <- read_csv(file = "data/pr_sdr/R_SDR_Run_Info.csv")

Light_Values <- unique(Run.Info$Light_Level)

# Join all coral and run metadata
metadata <- full_join(Sample.Info, Run.Info) %>%
  mutate(Date = as_date(as.character(Date), format = "%Y%m%d"))

starttime<-Run.Info$IntervalStart
endtime<-Run.Info$IntervalStop
```

```{r}
plate
Light_Values
date
```

Run loop to extract slopes from respiration data in each file name/plate. 

```{r, results=FALSE, warning=FALSE, message=FALSE}

for(file in 1:length(file.names)) { # for every file in list start at the first and run this following function
  
  #for (i in Light_Values) { #in every file, for each light value 
    
  Resp.Data <-read.table(file.path(path.p,file.names[file]), skip = 56, header=T, sep=",", na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Resp.Data$Temp <- Resp.Data[,30] #assigns temp column
  Resp.Data$Time.Min <- seq.int(from=0, to=((nrow(Resp.Data)*0.25)-0.25), by = 0.25) #set time in min
  
  #extract start and end times 
  starttime<-Run.Info%>%
    select(Plate, Light_Level, IntervalStart)%>%
    filter(Plate==plate[file])%>%
    #filter(Light_Level==i)%>%
    select(IntervalStart)%>%
    as.data.frame()

   starttime<-starttime[1,1]
   
   endtime<-Run.Info%>%
    select(Plate, Light_Level, IntervalStop)%>%
    filter(Plate==plate[file])%>%
    #filter(Light_Level==i)%>%
    select(IntervalStop)%>%
    as.data.frame()

   endtime<-endtime[1,1]
   
  #filter by light interval - need to bind in dataframe to tell it which start and end time to use 
  #Resp.Data <- Resp.Data %>% #filters data by interval for light 
     #filter(Time.Min > starttime)%>%
     #filter(Time.Min < endtime) 
  
  Resp.Data.N <- Resp.Data[,3:26] #subset desired columns
  
  #add column names back in 
  Resp.Data.N<-as.data.frame(Resp.Data.N)
  
  for(j in 1:(ncol(Resp.Data.N))){
    model <- rankLocReg(
      xall=Resp.Data$Time.Min, yall=as.numeric(Resp.Data.N[, j]), 
      alpha=0.4, method="pc", verbose=TRUE) #extract slopes, percentile rank method with minimum window size of 0.4. This means that in order to fit a slope, it has to encompass at least 40% of available datapoints. 
    
    pdf(paste0("output/P_regressions/",date[file], "_Plate",plate[file],"_",rename[j],"_light", Light_Values,"_regression_trunc.pdf")) #generate output file names
    plot(model)
    dev.off()
    
    Resp.Rb[j,1] <- as.character(date[file]) #stores the date
    Resp.Rb[j,2] <- as.character(plate[file]) #stores the run number
    #Resp.Rb[j,3] <- as.character(samp[j+(i-1)*ncol(Resp.Data.N)]) #stores the sample ID
    #Resp.Rb[j,3] <- as.character(samp[j])
    Resp.Rb[j,3] <- as.character(rename[j]) #stores the chamber ID
    Resp.Rb[j,4] <- as.character(1) #stores the chamber ID
    Resp.Rb[j,5:6] <- model$allRegs[1,c(4,5)] #inserts slope and intercept in the dataframe
    
  }
  
  Resp.R <- rbind(Resp.R, Resp.Rb) #bind final data frame
 # }
  
}

xx <- Resp.R 

Resp.R <- Resp.R[-1,] #remove empty row
plot(Resp.R$umol.L.min~as.factor(Resp.R$Plate), side = 2, las = 2, xlab="")

write.csv(Resp.R, paste0("output/resp_sdr_rates_raw.csv")) #save respiration rate data

```

Need to finalize this script
Calculate average temperature of each run and export to a table to confirm temperatures are not different between PAR values.    
```{r}
# list files 
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) 

#generate matrix to populate
Temp.R <- data.frame(matrix(NA, ncol=4))
colnames(Temp.R) <- c("Date", "Plate","Temp.C", "Assay")

Temp.Rb <- data.frame(matrix(NA, ncol=4))
colnames(Temp.Rb) <- c("Date", "Plate","Temp.C", "Assay")

#read in temps and generate mean values for each file  
for(file in 1:length(file.names)) { 
  
  #for(i in Light_Values) {# for every file in list start at the first and run this following function
  
    Temp.Data <-read.table(file.path(path.p,file.names[file]), skip = 56, header=T, sep=",", na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Temp.Data$Temp <- mean(Temp.Data[,30]) #assigns temp column

  
    Temp.Rb[j,1] <- as.character(date[file]) #stores the date
    Temp.Rb[j,2] <- as.character(plate[file]) #stores the run number
    Temp.Rb[j,3] <- Temp.Data$Temp[1] #stores the sample ID
    Temp.Rb[j,4] <- "Resp" #stores the assay name
    
    
    Temp.R <- rbind(Temp.R, Temp.Rb) #bind final data frame
    #Temp.R <- na.omit(Temp.R)
  
#}

}

Temp.R <- na.omit(Temp.R)

write.csv(Temp.R, paste0("output/resp_sdr_runs_temp.csv")) #save respiration rate data
```
