--
title: "Larval Photosynthesis Rates Extraction with LoLinR - Bermuda 2023 - PI Curves"
author: "Zoe Dellaert"
date: "2023-07-13"
output: html_document
---

This script is analyzing photosynthesis rates under different light levels for P. asteroides larvae in Bermuda. The code is based on A. Huffmyer's PI [script](https://github.com/AHuffmyer/larval_symbiont_TPC/blob/main/scripts/larval_pi_curve_sdr_analysis.Rmd). Important note - this data was collected in 4 mL vials, not the 24-well plates.

## Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('rTPC')
library('ggstatsplot')
library('nls.multstart')
library('broom')
```

# Data visualization and manipulation  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
pi_data<-read.csv("../output/larval_pi_curves/calculated_normalized_photo_rates.csv") #load data
```

Format data. 
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
pi_data<-pi_data[!is.na(pi_data$Type),]

#format columns
pi_data$SDR<-as.factor(pi_data$SDR)
pi_data$Plate<-as.factor(pi_data$Plate)
```

Look for outliers in the data.  

```{r}
boxplot(pi_data$P.nmol.org.min)

boxplot(pi_data$P.nmol.org.min~pi_data$PAR)
```

Replace any net negative value with 0, since this indicates no photosynthesis. 
```{r}
pi_data <-pi_data %>% mutate(P.nmol.org.min=if_else(P.nmol.org.min<0 & PAR>0, 0, P.nmol.org.min)) 
```

View data. Remove outliers. 
```{r}
boxplot(pi_data$P.nmol.org.min~pi_data$PAR)

pi_data <- pi_data%>%filter(P.nmol.org.min<0.06)

boxplot(pi_data$P.nmol.org.min)

```

Log transform the data.
```{r}
pi_data<-pi_data%>%
  mutate(P.nmol.org.min=log10(P.nmol.org.min+1))

boxplot(pi_data$P.nmol.org.min~pi_data$PAR)
```

Calculate mean temperature values for each run.    

```{r}
pi.temps<-read.csv("../output/larval_pi_curves/runs_temp.csv")
pi.temps = subset(pi.temps, select = -c(X) ) #remove empty column

pi_data$Plate<-as.integer(pi_data$Plate)
```


Add temperature data to master data frame.  
```{r}
pi_data<-left_join(pi_data,pi.temps)

#round to 0.1°C 
pi_data<-pi_data%>%
  mutate(Temp.C=round(Temp.C,1))
```


# Plot photosynthesis over PAR treatments  

Plot data with means   
```{r}
pi_plot1<-pi_data %>%
    group_by(PAR)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR))) +
    geom_hline(yintercept=0, linetype="dashed", color="black", linewidth=0.75)+
    geom_point(aes(group=interaction(PAR), colour=PAR), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    #scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_plot1

ggsave("../output/larval_pi_curves/figures/pi_means.png", pi_plot1, dpi=300, w=7, h=5, units="in")
```

Display with plate. 
```{r}
pi_plot1a<-pi_data %>%
    group_by(PAR, Plate)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR))) +
    facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR), colour=PAR), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    #scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      )

pi_plot1a
```

Plot data with a loess line   
```{r}
pi_plot2<-pi_data %>%
    
    ggplot(., aes(x = as.factor(PAR), y = P.nmol.org.min)) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_smooth(method="loess", alpha=0.4, se=FALSE)+
    xlab("Light (PAR)") + 
    #scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_plot2

ggsave("../output/larval_pi_curves/figures/pi_loess.png", pi_plot2, dpi=300, w=7, h=5, units="in")
```


# View respiration 

View by plate. 
```{r}
pi_plot3<-pi_data %>%
    group_by(PAR)%>%
    filter(PAR=="0")%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE)*-1, sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR))) +
    #facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR)), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    #scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      )
pi_plot3
```


Run anova on values 
```{r}
model<-aov(P.nmol.org.min~as.factor(PAR), data=pi_data)
summary(model)

library("emmeans")

emm<-emmeans(model, ~ PAR)
pairs(emm)
```

