---
title: "Larval Photosynthesis Rates Extraction with LoLinR - Moorea Pocillopora 2023 - PR Measurements"
author: "AS Huffmyer Hollie Putnam"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

Need to QC individual export plots 

## Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
library(devtools)
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr') 
if ("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc') 
if ("respR" %in% rownames(installed.packages()) == 'FALSE') install.packages('respR') 


#load packages
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('plyr')
library('dplyr')
library('stringr')
library('Rmisc')
library('respR')
library('tidyverse')

```

## Read in files  

Set the path of all respirometry files within the R project.   

```{r, warning=FALSE, message=FALSE}
path.p<-"data/pr_sdr/runP" #location of files
```

Bring in the file names.
```{r, warning=FALSE, message=FALSE}
# bring in the respiration file names
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) 
```

Bring in the run metadata. 
```{r}
run_metadata<-read.csv("data/pr_sdr/P_SDR_Run_Info.csv")
```

## Extract photosynthesis rates  

Generate photosynthesis data frames.

```{r, warning=FALSE, message=FALSE}
#generate a 7 column dataframe with specific column names
#photosynthesis
Photo.R <- data.frame(matrix(NA, ncol=7))
colnames(Photo.R) <- c("Date", "Plate","Sample.ID","Chamber.ID","Light", "Intercept", "umol.L.min")

Photo.Rb <- data.frame(matrix(NA, ncol=7))
colnames(Photo.Rb) <- c("Date", "Plate","Sample.ID","Chamber.ID","Light", "Intercept", "umol.L.min")

```

Load in the sample information file. It is important to have chambers in the order "A1, B1, C1, D1", rather than "A1, A2, A3, A4".  

Read in sample and run information and merge into a master metadata file. 
```{r}
# Load PI curve sample metadata (i.e., which larvae were in which runs)
Sample.Info <- read_csv(file = "data/pr_sdr/PR_SDR_Sample_Info.csv") #%>%
  #select(!Notes)

rename <- Sample.Info$Chamber_ID
samp <- Sample.Info$Sample_ID

plate <- as.numeric(str_sub(file.names, 26, 27))
date <- str_sub(file.names, 4, str_length(file.names)-26) #grab date from file name

# Load PI curve run metadata (i.e., light levels and interval times for each run)
Run.Info <- read_csv(file = "data/pr_sdr/P_SDR_Run_Info.csv")

Light_Values <- unique(Run.Info$Light_Level)

# Join all coral and run metadata
metadata <- full_join(Sample.Info, Run.Info) %>%
  mutate(Date = as_date(as.character(Date), format = "%Y%m%d"))

starttime<-Run.Info$IntervalStart
endtime<-Run.Info$IntervalStop
```

```{r}
plate
Light_Values
date
```

Run loop to extract slopes from photosynthesis data from each light level in each file name/plate. 

```{r, results=FALSE, warning=FALSE, message=FALSE}

for(file in 1:length(file.names)) { # for every file in list start at the first and run this following function
  
  for (i in Light_Values) { #in every file, for each light value 
    
  Photo.Data <-read.table(file.path(path.p,file.names[file]), skip = 56, header=T, sep=",", na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Photo.Data$Temp <- Photo.Data[,20] #assigns temp column
  Photo.Data$Time.Min <- seq.int(from=0, to=((nrow(Photo.Data)*0.25)-0.25), by = 0.25) #set time in min
  
  #extract start and end times 
  starttime<-Run.Info%>%
    select(Plate, Light_Level, IntervalStart)%>%
    filter(Plate==plate[file])%>%
    filter(Light_Level==i)%>%
    select(IntervalStart)%>%
    as.data.frame()

   starttime<-starttime[1,1]
   
   endtime<-Run.Info%>%
    select(Plate, Light_Level, IntervalStop)%>%
    filter(Plate==plate[file])%>%
    filter(Light_Level==i)%>%
    select(IntervalStop)%>%
    as.data.frame()

   endtime<-endtime[1,1]
   
  #filter by light interval - need to bind in dataframe to tell it which start and end time to use 
  Photo.Data <- Photo.Data %>% #filters data by interval for light 
     filter(Time.Min > starttime)%>%
     filter(Time.Min < endtime) 
  
  Photo.Data.N <- Photo.Data[,3:15] #subset desired columns
  
  #add column names back in 
  Photo.Data.N<-as.data.frame(Photo.Data.N)
  
  for(j in 1:(ncol(Photo.Data.N))){
    model <- rankLocReg(
      xall=Photo.Data$Time.Min, yall=as.numeric(Photo.Data.N[, j]), 
      alpha=0.4, method="pc", verbose=TRUE) #extract slopes, percentile rank method with minimum window size of 0.4. This means that in order to fit a slope, it has to encompass at least 40% of available datapoints. 
    
    pdf(paste0("output/P_regressions/",date[file], "_Plate",plate[file],"_",rename[j],"_light", Light_Values[i],"_regression_trunc.pdf")) #generate output file names
    plot(model)
    dev.off()
    
    Photo.Rb[j,1] <- as.character(date[file]) #stores the date
    Photo.Rb[j,2] <- as.character(plate[file]) #stores the run number
    #Photo.Rb[j,3] <- as.character(samp[j+(i-1)*ncol(Photo.Data.N)]) #stores the sample ID
    Photo.Rb[j,3] <- as.character(samp[j])
    Photo.Rb[j,4] <- as.character(rename[j]) #stores the chamber ID
    Photo.Rb[j,5] <- as.character(i) #stores the chamber ID
    Photo.Rb[j,6:7] <- model$allRegs[i,c(4,5)] #inserts slope and intercept in the dataframe
    
  }
  
  Photo.R <- rbind(Photo.R, Photo.Rb) #bind final data frame
  }
  
}

xx <- Photo.R 

Photo.R <- Photo.R[-1,] #remove empty row
plot(Photo.R$umol.L.min~as.factor(Photo.R$Plate), side = 2, las = 2, xlab="")

write.csv(Photo.R, paste0("output/photo_sdr_rates_raw.csv")) #save respiration rate data

```

Calculate average temperature of each run and export to a table to confirm temperatures are not different between PAR values.    
```{r}
# list files 
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) 

#generate matrix to populate
Temp.P <- data.frame(matrix(NA, ncol=4))
colnames(Temp.P) <- c("Date", "Plate","Temp.C", "Light")

Temp.Pb <- data.frame(matrix(NA, ncol=4))
colnames(Temp.Pb) <- c("Date", "Plate","Temp.C", "Light")

#read in temps and generate mean values for each step of the light curve  
for(file in 1:length(file.names)) { 
  
  for(i in Light_Values) {# for every file in list start at the first and run this following function
  
    Temp.Data <-read.table(file.path(path.p,file.names[file]), skip = 56, header=T, sep=",", na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Temp.Data$Temp <- Temp.Data[,20] #assigns temp column
  Temp.Data$Time.Min <- seq.int(from=0, to=((nrow(Temp.Data)*0.25)-0.25), by = 0.25) #set time in min
  
  #extract start and end times 
  starttime<-Run.Info%>%
    select(Plate, Light_Level, IntervalStart)%>%
    filter(Plate==plate[file])%>%
    filter(Light_Level==i)%>%
    select(IntervalStart)%>%
    as.data.frame()

   starttime<-starttime[1,1]
   
   endtime<-Run.Info%>%
    select(Plate, Light_Level, IntervalStop)%>%
    filter(Plate==plate[file])%>%
    filter(Light_Level==i)%>%
    select(IntervalStop)%>%
    as.data.frame()

   endtime<-endtime[1,1]
  
  #filter by light interval - need to bind in dataframe to tell it which start and end time to use 
  Temp.Data <- Temp.Data %>% #filters data by interval for light 
     filter(Time.Min > starttime)%>%
     filter(Time.Min < endtime) 
  
    Temp.Pb[j,1] <- as.character(date[file]) #stores the date
    Temp.Pb[j,2] <- as.character(plate[file]) #stores the run number
    Temp.Pb[j,3] <- mean(Temp.Data$Temp) #stores the sample ID
    Temp.Pb[j,4] <- mean(Light_Values[i]) #stores the sample ID
    
    
    Temp.P <- rbind(Temp.P, Temp.Pb) #bind final data frame
    Temp.P <- na.omit(Temp.P)
  
}

}
write.csv(Temp.P, paste0("output/photo_sdr_runs_temp.csv")) #save respiration rate data
```

Save photosynthesis data frames.  

```{r, results=FALSE, warning=FALSE, message=FALSE}



```





