---
title: "Analyzing larval environmental data - Hawaii 2023"
author: "Ariana S Huffmyer"
date: "2023"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

This script reads and plots environmental data from Hobo tidbit loggers and daily measurements.  

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
```

# Tidbit analysis 

## Read in files 

Read in Hobo Tidbit files  
```{r}
tidbit.files <- list.files(path="data/environmental/tidbit_loggers", pattern="*.csv", full.names=TRUE)

#will need to change to HST time 
tidbits <- setNames(tidbit.files, tidbit.files) %>%
    map_dfr(read_csv, .id = "logger") %>%
    rename(DateTime=`Date-Time (HST)`, temp.C=`Ch: 1 - Temperature   (°C)`)%>%
    select(!`#`)%>%
    select(logger, DateTime, temp.C)%>%
    mutate(logger=sub(".*/", "", logger))%>% #remove the path name
    mutate(logger=sub("_.*", "", logger))%>% #keep only the serial number 
    filter(!temp.C=="NA")

tidbits$DateTime<-as.POSIXct(tidbits$DateTime, format="%m/%d/%Y %H:%M:%S", tz="Pacific/Honolulu")
```

Remove time periods during logger read out.   
```{r}
#tidbits<-tidbits%>%filter(DateTime<as.POSIXct('2022-10-28 16:30:00', tz="Pacific/Honolulu"))
```

List the logger serial numbers in this dataframe.  
```{r}
levels(as.factor(tidbits$logger))
```

There are 6 loggers as expected.  

## Read in treatment information  

Read in metadata that assigns a pendant serial number to a tank number and/or treatment.  
```{r}
metadata<-read_csv("data/environmental/logger_metadata.csv")
metadata$logger<-as.character(metadata$logger)
```

Assign treatment information to logger data.  
```{r}
tidbits<-left_join(tidbits, metadata, by="logger")
tidbits$tank<-as.factor(tidbits$tank)
tidbits$group<-as.factor(tidbits$group)
tidbits$project<-as.factor(tidbits$project)
tidbits$type<-as.factor(tidbits$type)
```

# Developmental time series project 

Plot temperature by incubator colored by treatment. 

```{r}
temp_plot1<-tidbits%>%
  filter(project=="dev_timeseries")%>%
  filter(DateTime < as.POSIXct("2023-06-20 21:00:00"))%>%
  filter(DateTime > as.POSIXct("2023-06-16 21:00:00"))%>%
  
  ggplot(aes(x=DateTime, y=temp.C, colour=group))+
  geom_point()+
  #ylim(26, 30)+
  scale_colour_manual(values=c("blue", "red"), name="Temperature")+
  ylab("Temperature (°C)")+
  xlab("Date Time")+
  theme_classic(); temp_plot1
```

Summarize by treatment with error shading for standard error.  
```{r}
temp_plot_treatment1<-tidbits%>%
  filter(project=="dev_timeseries")%>%
  filter(DateTime < as.POSIXct("2023-06-20 21:00:00"))%>%
  filter(DateTime > as.POSIXct("2023-06-17 23:00:00"))%>%
  group_by(group, DateTime)%>%
  summarise(mean=mean(temp.C, na.rm=TRUE), sd=sd(temp.C, na.rm=TRUE), sem=sd/sqrt(6), upper=mean+sd, lower=mean-sd)%>%
  
  ggplot(aes(x=DateTime, y=mean, colour=group, fill=group))+
  #geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper, colour=group), alpha = 0.4, colour = NA, show.legend=FALSE) +
  geom_smooth(method="loess", span=0.1, se=FALSE, aes(colour=group), alpha=1, show.legend=TRUE) +
  geom_vline(xintercept=as.POSIXct("2023-06-17 23:30:00"), linetype="dashed", color="black")+
  geom_vline(xintercept=as.POSIXct("2023-06-18 02:45:00"), linetype="dashed", color="black")+
  geom_vline(xintercept=as.POSIXct("2023-06-18 02:45:00"), linetype="dashed", color="black")+
  geom_vline(xintercept=as.POSIXct("2023-06-18 09:00:00"), linetype="dashed", color="black")+
  geom_vline(xintercept=as.POSIXct("2023-06-18 14:00:00"), linetype="dashed", color="black")+
  geom_vline(xintercept=as.POSIXct("2023-06-18 21:00:00"), linetype="dashed", color="black")+
  geom_vline(xintercept=as.POSIXct("2023-06-19 02:30:00"), linetype="dashed", color="black")+
  geom_vline(xintercept=as.POSIXct("2023-06-19 21:00:00"), linetype="dashed", color="black")+
  geom_vline(xintercept=as.POSIXct("2023-06-20 21:00:00"), linetype="dashed", color="black")+
  #ylim(26, 30)+
  ylab("Temperature (°C)")+
  xlab("Date Time")+
  scale_colour_manual(values=c("blue", "red"), name="Temperature")+
  scale_fill_manual(values=c("blue", "red"), name="Temperature")+
  theme_classic()+
  theme(
    axis.text=element_text(color="black", size=11),
    axis.title=element_text(color="black", size=14),
    legend.position="none"
  ); temp_plot_treatment1

ggsave("figures/environmental/dev_timeseries/tidbit_temperature_smooth.png", temp_plot_treatment1, width=6, height=4)
```

Calculate the mean difference in temperature treatments between groups.  
```{r}
tidbits%>%
  filter(project=="dev_timeseries")%>%
  filter(DateTime < as.POSIXct("2023-06-20 21:00:00"))%>%
  filter(DateTime > as.POSIXct("2023-06-16 21:00:00"))%>%
  group_by(group, DateTime)%>%
  summarise(mean=mean(temp.C, na.rm=TRUE))%>%
  spread(key=group, value=mean)%>%
  reframe(difference=high-ambient)%>%
  reframe(mean=mean(difference))
```

Calculate mean temperature for each tank.      
```{r}
tidbits %>%
  filter(project=="dev_timeseries")%>%
  filter(DateTime < as.POSIXct("2023-06-20 21:00:00"))%>%
  filter(DateTime > as.POSIXct("2023-06-16 21:00:00"))%>%
  mutate(day = floor_date(DateTime, "day")) %>%
           group_by(tank, group) %>%
           summarise(mean = mean(temp.C))
```


# Larval symbiont project 

Plot temperature by incubator colored by treatment. 

```{r}
temp_plot2a<-tidbits%>%
  filter(project=="larvae_cd")%>%
  filter(tank %in% c("ambient_tank", "moderate_tank", "high_tank"))%>%
  filter(DateTime < as.POSIXct("2023-06-27 15:00:00"))%>%
  filter(DateTime > as.POSIXct("2023-06-24 13:00:00"))%>%
  filter(!c(DateTime > as.POSIXct("2023-06-24 19:45:00") & DateTime < as.POSIXct("2023-06-27 8:00:00")))%>%
  
  filter(DateTime < as.POSIXct("2023-06-24 19:45:00"))%>%
  
  ggplot(aes(x=DateTime, y=temp.C, colour=group))+
  geom_point()+
  scale_colour_manual(values=c("lightblue", "blue", "red"), name="Temperature")+
  ylab("Temperature (°C)")+
  xlab("Time")+
  ylim(26,34)+
  ggtitle("2023-06-24")+
  theme_classic()+
  theme(
    axis.text=element_text(color="black", size=11),
    axis.title=element_text(color="black", size=14),
    legend.position="none"
  ); temp_plot2a
```

```{r}
temp_plot2b<-tidbits%>%
  filter(project=="larvae_cd")%>%
  filter(tank %in% c("ambient_tank", "moderate_tank", "high_tank"))%>%
  filter(DateTime < as.POSIXct("2023-06-27 15:00:00"))%>%
  filter(DateTime > as.POSIXct("2023-06-24 13:00:00"))%>%
  filter(!c(DateTime > as.POSIXct("2023-06-24 19:45:00") & DateTime < as.POSIXct("2023-06-27 8:00:00")))%>%
  
  filter(DateTime > as.POSIXct("2023-06-27 08:00:00"))%>%
  
  ggplot(aes(x=DateTime, y=temp.C, colour=group))+
  geom_point()+
  scale_colour_manual(values=c("lightblue", "blue", "red"), name="Temperature")+
  ylab("Temperature (°C)")+
  xlab("Time")+
  ylim(26,34)+
  ggtitle("2023-06-27")+
  theme_classic()+
  theme(
    axis.text=element_text(color="black", size=11),
    axis.title=element_text(color="black", size=14),
    legend.position="none"
  ); temp_plot2b
```

Combine the plots from each day. 
```{r}
incubation_plots<-plot_grid(temp_plot2a, temp_plot2b, align="vh")

ggsave("figures/environmental/incubation_temps.png", incubation_plots, width=8, height=4, dpi=300)
```

Calculate mean temperature for each treatment      
```{r}
tidbits %>%
  filter(project=="larvae_cd")%>%
  filter(tank %in% c("ambient_tank", "moderate_tank", "high_tank"))%>%
  filter(DateTime < as.POSIXct("2023-06-27 15:00:00"))%>%
  filter(DateTime > as.POSIXct("2023-06-24 13:00:00"))%>%
  filter(!c(DateTime > as.POSIXct("2023-06-24 19:45:00") & DateTime < as.POSIXct("2023-06-27 8:00:00")))%>%
  
  mutate(day = floor_date(DateTime, "day")) %>%
           group_by(tank, group) %>%
           summarise(mean = mean(temp.C))#%>%
         # group_by(group) %>%
          # summarise(mean=mean(mean))
```
