---
title: "Logger analysis"
author: "Jill Ashey"
date: "2023-08-18"
output: html_document
---

This script reads and plots temperature data from Hobo loggers from the PAST experiment in Bermuda 2023. 

## Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

Load packages
```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
```

## Read in files 

Read in Hobo files 
```{r}
# Get list of csv files
file_paths <- list.files("data/enviromental/HoboLoggerData/loggers/", pattern = "\\.csv$", full.names = TRUE)

# Initialize an empty list to store processed data frames
processed_data <- list()

# Process each file and store in list 
for (file_path in file_paths) {
  data <- read_csv(file_path, skip = 1) %>% # read in each file, skip first row
    select(2:4) %>% # select columns 2 through 4
    rename(Date.Time = 1, temp_c = 2, light_lux = 3) %>% # rename columns
    mutate(logger = sub("^(.*?)_.+", "\\1", basename(file_path)), # Extract logger name and add as new column
           DateEnded = sub(".+_(.*)$", "\\1", basename(file_path))) # Extract DateEnded and add as new column
  processed_data[[file_path]] <- data # save data into list
}

# Bind data into one df
final_df <- bind_rows(processed_data)

# Remove .csv from the end of DateEnded
final_df$DateEnded <- gsub(".csv", "", final_df$DateEnded)
```

Set time format
```{r}
final_df$Date.Time <- as.POSIXct(final_df$Date.Time, format="%m/%d/%Y %I:%M:%S %p")
```

Remove periods during logger read out 

List logger serial numbers in this dataframe 
```{r}
levels(as.factor(final_df$logger))
```

## Read in treatment information  

Read in metadata that assigns a pendant serial number to a tank number and/or treatment.  
```{r}
metadata<-read_csv("data/enviromental/Logger_metadata.csv")

# Rename LoggerSN as logger
metadata <- metadata %>%
  rename(logger = LoggerSN)

# Set logger and DateEndedas a character
metadata$logger<-as.character(metadata$logger)
metadata$DateEnded <- as.character(metadata$DateEnded)
```

Join metadata and final_df files by logger and Date Ended
```{r}
loggers <- left_join(final_df, metadata, by = c("logger", "DateEnded")) %>%
  drop_na(temp_c)
```

## Plot logger data 

Calculate mean temperature for each tank.      
```{r}
means <- loggers %>%
  group_by(DateEnded, Type, Treatment, Basin) %>%
  summarise(mean = mean(temp_c),
            sd = sd(temp_c))
```

Plot by treatment
```{r}
plot1 <- ggplot(loggers, aes(x=Date.Time, y=temp_c, colour=Treatment))+
  geom_point()+
  geom_vline(xintercept=as.POSIXct("0023-07-21 23:00:00"), linetype="dashed", color="black")+
  #ylim(26, 30)+
  scale_colour_manual(values=c("blue", "red"), name="Temperature")+
  ylab("Temperature (°C)")+
  xlab("Date Time")+
  theme_classic(); plot1

ggsave(filename="output/environmental/loggers/logger_by_treatment.png", plot=plot1, dpi=300, width=9, height=6, units="in")
```

Plot by logger location
```{r}
plot2 <- ggplot(loggers, aes(x=Date.Time, y=temp_c, colour=LoggerID))+
  geom_point()+
  #ylim(26, 30)+
  #scale_colour_manual(values=c("blue", "red"), name="Temperature")+
  ylab("Temperature (°C)")+
  xlab("Date Time")+
  theme_classic(); plot2

ggsave(filename="output/environmental/logger_by_ID.png", plot=plot2, dpi=300, width=9, height=6, units="in")
```

Add vertical lines to plot so we know when sampling occurred. Line type will be release, larval sampling, spat sampling, etc. Line color will be cohort (ie which date did colonies release on)

