---
title: "PAM data analysis"
author: "Jill Ashey"
date: "2023-08-14"
output: html_document
---

This code analyzes the pulse amplitude fluometry (PAM) data that was collected from P. asteriodes larvae and spat in Bermuda 2023. 

## Set up

```{r setup, include=FALSE}
if ("knitr" %in% rownames(installed.packages()) == 'FALSE') install.packages('knitr') 
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("RColorBrewer" %in% rownames(installed.packages()) == 'FALSE') install.packages('RColorBrewer') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("knitr" %in% rownames(installed.packages()) == 'FALSE') install.packages('knitr') 
if ("kableExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('kableExtra') 
if ("GGally" %in% rownames(installed.packages()) == 'FALSE') install.packages('GGally') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("Hmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Hmisc') 
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggstatsplot') 

library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(car)
library(emmeans)
library(multcomp)
library(knitr)
library(kableExtra)
library(lme4)
library(GGally)
library(lmerTest)
library(cowplot)
library(Hmisc)
library(ggstatsplot)
```

## First, I am going to analyze the data from the larval samples. 

## Load data 
```{r}
pam_data <- read_csv("../data/pam/pam_larvae_dark.csv")
head(pam_data)
tail(pam_data)
```

I updated the data so that the pam larvae csv included DateCollected, DateMeasured, TimeMeasured, Colony, Basin, and Squarical. Now we can do plotting and stats on this data!

## Manipulate data frame 

Remove space from condition and colony columns
```{r}
pam_data$Colony <- sub(" ", "", pam_data$Colony)
```

Remove NAs from colony 
```{r}
pam_data <- pam_data %>%
  drop_na(Colony)
```

Set date
```{r}
pam_data$DateCollected <- as.POSIXct(pam_data$DateCollected, format = "%m/%d/%y")
pam_data$DateMeasured <- as.POSIXct(pam_data$DateMeasured, format = "%m/%d/%y")
```

## Make plots, identify outliers and do statistics 

### Fv.Fm

Plot Fv/Fm by treatment, colony and date
```{r}
ggplot(pam_data, aes(x = Colony, y = Fv.Fm, color = Treatment)) +
  geom_boxplot(aes(colour=Treatment,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw() +
  facet_wrap(~DateMeasured)
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

Some of the data only have one point...We may not be able to use these measurements if the measurement is n=1. 

NEED TO LOOK AT OUTLIERS 

Do stats on Fv/Fm by treatment, colony and date 
```{r}
model <- aov(Fv.Fm ~ condition*colony*date, data = pam_data_spat)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ condition*colony*date, data = pam_data_spat)
# anova assumptions NOT met  

# View post-hoc differences 
emm = emmeans(model, ~ Treatment*Colony*DateMeasured)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)

# confusion
```

Plot Fv/Fm by treatment and colony
```{r}
ggplot(pam_data, aes(x = Colony, y = Fv.Fm)) +
  geom_boxplot(aes(colour=Treatment), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw()
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

Some of the data only have one point...We may not be able to use these measurements if the measurement is n=1. 

Do stats on Fv/Fm by treatment and colony 
```{r}
model <- aov(Fv.Fm ~ Treatment*Colony, data = pam_data)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ Treatment*Colony*DateMeasured, data = pam_data)
# anova assumptions met 

# View post-hoc differences 
TukeyHSD(model)

emm = emmeans(model, ~ Treatment*Colony)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)
```

Plot by treatment only
```{r}
ggplot(pam_data, aes(x = Treatment, y = Fv.Fm)) +
  geom_boxplot(aes(colour=Treatment,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw() 
  #facet_wrap(~date)
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

Do stats on Fv/Fm by treatment 
```{r}
model <- aov(Fv.Fm ~ Treatment, data = pam_data)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ Treatment*Colony*DateMeasured, data = pam_data)
# anova assumptions met 

# View post-hoc differences 
TukeyHSD(model)

emm = emmeans(model, ~ Treatment*Colony)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)
```

## Count observations 

Calculate the number of measurements grouped by date, treatment and colony
```{r}
summary_data <- pam_data %>%
  group_by(DateMeasured, Colony, Treatment) %>%
  summarise(observations = n())

summary_data <- pam_data %>%
  group_by(Colony, Treatment) %>%
  summarise(observations = n())
```

We may only be able to use the measurements with >3 data points per date, treatment and colony if that makes sense. 

Group by colony and plot so that date is the x-axis, Fv/Fm is the y-axis, colors are treatment, and shapes are colony
```{r}
test <- pam_data %>%
  group_by(Colony, DateMeasured, Treatment) %>%
  summarise(mean = mean(Fv.Fm),
            sd = sd(Fv.Fm)) %>%
  drop_na(sd)

# Plot
ggplot(test, aes(x=DateMeasured, y=mean, color = Treatment, shape = Colony)) +
  #geom_line() +
  #geom_boxplot() +
  geom_point(size = 5) + 
  scale_shape_manual(values = c(0:15))+ 
  scale_color_manual(values = c("blue", "red")) +
  theme_classic()
```


































## Now, I am going to analyze the data from the spat samples. 

## Load data 
```{r}
pam_data_spat <- read.csv("data/pam/pam_spat_measurements.csv", na.strings=c("","NA"))
head(pam_data_spat)
tail(pam_data_spat)
```

## Manipulate data frame 

Remove space from condition and colony columns
```{r}
pam_data_spat$condition <- sub(" ", "", pam_data_spat$condition)
pam_data_spat$colony <- sub(" ", "", pam_data_spat$colony)
```

Remove NAs from tank and colony 
```{r}
pam_data_spat <- pam_data_spat %>%
  drop_na(condition, colony)
```
No data points removed here. 

Set date
```{r}
pam_data_spat$date <- as.POSIXct(pam_data_spat$date, format = "%m/%d/%y")
```

## Make plots, identify outliers and do statistics 

### Fv.Fm

Plot Fv/Fm by treatment, colony and date
```{r}
ggplot(pam_data_spat, aes(x = colony, y = Fv.Fm, color = condition)) +
  geom_boxplot(aes(colour=condition,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw() +
  facet_wrap(~date)
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

Important note!! Spat samples that were supposed to be PAMmed on July 29 were actually PAMmed on July 28. Team took the day off on July 29. 

NEED TO LOOK AT OUTLIERS STILL 

Not sure how to do the ANOVA for this... maybe 3 way anova?? 
Do stats - run a two-way ANOVA
```{r}
model <- aov(Fv.Fm ~ condition*colony*date, data = pam_data_spat)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ condition*colony*date, data = pam_data_spat)
```

There is a violation of variance assumptions. Conduct non-parametric test 
```{r}
kruskal.test(Fv.Fm ~ colony, data = pam_data_spat)
kruskal.test(Fv.Fm ~ condition, data = pam_data_spat)
kruskal.test(Fv.Fm ~ date, data = pam_data_spat)

# View post-hoc differences 
emm = emmeans(model, ~ condition*colony*date)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)
```

Plot Fv/Fm by treatment and colony
```{r}
ggplot(pam_data_spat, aes(x = colony, y = Fv.Fm)) +
  geom_boxplot(aes(colour=condition), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw()
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

LOOK AT OUTLIERS IN DATA 

Do stats on Fv/Fm by treatment and colony 
```{r}
model <- aov(Fv.Fm ~ condition*colony, data = pam_data_spat)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ condition*colony, data = pam_data_spat)
# anova assumptions NOT met 
```

There is a violation of variance assumptions. Conduct non-parametric test 
```{r}
# View post-hoc differences 
emm = emmeans(model, ~ condition*colony)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)
```
























 


## Now, I am going to analyze the data from the spat samples. 

## Load data 
```{r}
pam_data <- read.csv("data/pam/pam_spat_measurements.csv", na.strings=c("","NA"))
head(pam_data)
tail(pam_data)
```

First look at the data - don't have info about their 'chamber' or tank id. 

## Manipulate data frame 

Remove space from condition and colony columns
```{r}
pam_data$condition <- sub(" ", "", pam_data$condition)
pam_data$colony <- sub(" ", "", pam_data$colony)
```

Remove NAs from tank and colony 
```{r}
pam_data <- pam_data %>%
  drop_na(condition, colony)
```
No data points removed here. 

Set date
```{r}
pam_data$date <- as.POSIXct(pam_data$date, format = "%m/%d/%y")
```

## Make plots, identify outliers and do statistics 

### Fv.Fm

Plot Fv/Fm
```{r}
ggplot(pam_data, aes(x = colony, y = Fv.Fm, color = condition)) +
  geom_boxplot(aes(colour=condition,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw() +
  facet_wrap(~date)
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

NEED TO LOOK AT OUTLIERS STILL 

Not sure how to do the ANOVA for this... maybe 3 way anova?? 
Do stats - run a two-way ANOVA
```{r}
model <- aov(Fv.Fm ~ condition*colony*date, data = pam_data)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ condition*colony*date, data = pam_data)
```

There is a violation of variance assumptions. Conduct non-parametric test 
```{r}
kruskal.test(Fv.Fm ~ colony, data = pam_data)
kruskal.test(Fv.Fm ~ condition, data = pam_data)
kruskal.test(Fv.Fm ~ date, data = pam_data)

# View post-hoc differences 
emm = emmeans(model, ~ condition*colony*date)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)
```

What if I plotted Fv.Fm for spat without the date? 
```{r}
ggplot(pam_data, aes(x = colony, y = Fv.Fm, color = condition)) +
  geom_boxplot(aes(colour=condition,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw() 
  #facet_wrap(~date)
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()

ggplot(pam_data, aes(x = condition, y = Fv.Fm)) +
  geom_boxplot(aes(colour=condition,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw() 
  #facet_wrap(~date)
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

```{r}
model <- aov(Fv.Fm ~ condition*colony, data = pam_data)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ condition*colony*date, data = pam_data)
```

```{r}
t.test(Fv.Fm ~ condition, data = pam_data)
```







