---
title: "PAM data analysis"
author: "Jill Ashey"
date: "2023-08-14"
output: html_document
---

This code analyzes the pulse amplitude fluometry (PAM) data that was collected from P. asteriodes larvae and spat in Bermuda 2023. 

## Set up

```{r setup, include=FALSE}
if ("knitr" %in% rownames(installed.packages()) == 'FALSE') install.packages('knitr') 
knitr::opts_chunk$set(echo = FALSE)
```

Load packages
```{r}
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("RColorBrewer" %in% rownames(installed.packages()) == 'FALSE') install.packages('RColorBrewer') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("knitr" %in% rownames(installed.packages()) == 'FALSE') install.packages('knitr') 
if ("kableExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('kableExtra') 
if ("GGally" %in% rownames(installed.packages()) == 'FALSE') install.packages('GGally') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("Hmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Hmisc') 
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggstatsplot') 

library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(car)
library(emmeans)
library(multcomp)
library(knitr)
library(kableExtra)
library(lme4)
library(GGally)
library(lmerTest)
library(cowplot)
library(Hmisc)
library(ggstatsplot)
```

## First, I am going to analyze the data from the larval samples. 

## Load data 
```{r}
pam_data <- read_csv("data/pam/pam_larvae_dark.csv")
head(pam_data)
tail(pam_data)
```

I updated the data so that the pam larvae csv included DateCollected, DateMeasured, TimeMeasured, Colony, Basin, and Squarical. Now we can do plotting and stats on this data!

## Manipulate data frame 

Remove space from condition and colony columns
```{r}
pam_data$Colony <- sub(" ", "", pam_data$Colony)
```

Remove NAs from colony 
```{r}
pam_data <- pam_data %>%
  drop_na(Colony)

pam_data <- pam_data %>%
  filter(!Colony=="Pooled")

```

Set date
```{r}
pam_data$DateCollected <- as.POSIXct(pam_data$DateCollected, format = "%m/%d/%y")
#pam_data$DateMeasured <- as.POSIXct(pam_data$DateMeasured, format = "%m/%d/%y")
```

## Make plots, identify outliers and do statistics 

### Fv.Fm

Plot Fv/Fm by treatment and date
```{r}

pdf(file="output/larval.FvFm.treatment.date.pdf", width=8, height=4)

ggplot(pam_data, aes(x = DateMeasured, y = Fv.Fm, color = Treatment)) +
  geom_boxplot(aes(colour=Treatment,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  scale_colour_manual(values=c("blue", "red"))+
  ylab("Fv/Fm")+
  xlab("Date Measured")+
  theme_bw() 

dev.off()
```
Plot Fv/Fm by treatment, colony
```{r}

pdf(file="output/larval.FvFm.treatment.colony.pdf", width=8, height=4)


ggplot(pam_data, aes(x = Colony, y = Fv.Fm, color = Treatment)) +
  geom_boxplot(aes(colour=Treatment,), width=.5, outlier.shape= NA) +
  scale_colour_manual(values=c("blue", "red"))+
  ylab("Fv/Fm")+
  xlab("Parent Colony")+
  theme_bw() 

dev.off()
```

Plot Fv/Fm by treatment, colony, and date
```{r}

pdf(file="output/larval.FvFm.treatment.date.colony.pdf", width=8, height=4)

ggplot(pam_data, aes(x = Colony, y = Fv.Fm, color = Treatment)) +
  geom_boxplot(aes(colour=Treatment,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  scale_colour_manual(values=c("blue", "red"))+
  ylab("Fv/Fm")+
  xlab("Parent Colony")+
  facet_wrap("DateMeasured")+
  theme_bw() 

dev.off()
```

Plot Fv/Fm by treatment
```{r}

pdf(file="output/larval.FvFm.treatment.pdf", width=4, height=4)

ggplot(pam_data, aes(x = Treatment, y = Fv.Fm, color = Treatment)) +
  geom_boxplot(aes(color=Treatment,), width=.5, outlier.shape= NA) +
  geom_point(pch = 16, size = 1) +
  scale_colour_manual(values=c("blue", "red"))+
  ylab("Fv/Fm")+
  theme_bw() 

dev.off()

```

Some of the data only have one point...We may not be able to use these measurements if the measurement is n=1. 

Do stats on Fv/Fm by treatment, colony and date 
```{r}
model <- aov(Fv.Fm ~ Treatment*Colony*DateMeasured, data = pam_data)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ Treatment*Colony*DateMeasured, data = pam_data)
# anova assumptions  met  

# View post-hoc differences 
#emm = emmeans(model, ~ Treatment*Colony*DateMeasured)
#cld(emm, Letters=c(LETTERS)) #letter display
#pairs(emm)


```


## Count observations 

Calculate the number of measurements grouped by date, treatment and colony
```{r}
pam_data %>%
  group_by(DateMeasured, Colony, Treatment) %>%
  summarise(observations = n())

pam_data %>%
  group_by(Colony, Treatment) %>%
  summarise(observations = n())
```



## Now, I am going to analyze the data from the spat samples. 

## Load data 
```{r}
pam_data_spat <- read.csv("data/pam/pam_spat_measurements.csv", na.strings=c("","NA"))
head(pam_data_spat)
tail(pam_data_spat)
```

## Manipulate data frame 

Remove space from condition and colony columns
```{r}
pam_data_spat$condition <- sub(" ", "", pam_data_spat$condition)
pam_data_spat$colony <- sub(" ", "", pam_data_spat$colony)
```

Remove NAs from tank and colony 
```{r}
pam_data_spat <- pam_data_spat %>%
  drop_na(condition, colony)

pam_data_spat <- pam_data_spat %>%
  filter(!colony=="Pooled")
```
No data points removed here. 

Set date
```{r}
#pam_data_spat$date <- as.POSIXct(pam_data_spat$date, format = "%m/%d/%y")
```

## Make plots, identify outliers and do statistics 

### Fv.Fm

Plot Fv/Fm by treatment and date
```{r}

pdf(file="output/spat.FvFm.treatment.date.pdf", width=8, height=4)

ggplot(pam_data_spat, aes(x = date, y = Fv.Fm, color = condition)) +
  geom_boxplot(aes(colour=condition,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  scale_colour_manual(values=c("blue", "red"))+
  ylab("Fv/Fm")+
  xlab("Date Measured")+
  theme_bw() 

dev.off()
```
Plot Fv/Fm by treatment, colony
```{r}

pdf(file="output/spat.FvFm.treatment.colony.pdf", width=8, height=4)


ggplot(pam_data_spat, aes(x = colony, y = Fv.Fm, color = condition)) +
  geom_boxplot(aes(colour=condition,), width=.5, outlier.shape= NA) +
  scale_colour_manual(values=c("blue", "red"))+
  ylab("Fv/Fm")+
  xlab("Parent Colony")+
  theme_bw() 

dev.off()
```

Plot Fv/Fm by treatment, colony, and date
```{r}

pdf(file="output/spat.FvFm.treatment.date.colony.pdf", width=8, height=4)

ggplot(pam_data_spat, aes(x = colony, y = Fv.Fm, color = condition)) +
  geom_boxplot(aes(colour=condition,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  scale_colour_manual(values=c("blue", "red"))+
  ylab("Fv/Fm")+
  xlab("Parent Colony")+
  facet_wrap("date")+
  theme_bw() 

dev.off()
```

Plot Fv/Fm by treatment
```{r}

pdf(file="output/spat.FvFm.treatment.pdf", width=4, height=4)

ggplot(pam_data_spat, aes(x = condition, y = Fv.Fm, color = condition)) +
  geom_boxplot(aes(color=condition,), width=.5, outlier.shape= NA) +
  geom_point(pch = 16, size = 1) +
  scale_colour_manual(values=c("blue", "red"))+
  ylab("Fv/Fm")+
  theme_bw() 

dev.off()

```


Do stats on Fv/Fm by treatment, colony and date 
```{r}
model <- aov(Fv.Fm ~ condition*colony*date, data = pam_data_spat)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ condition*colony*date, data = pam_data_spat)
# anova assumptions  met  

# View post-hoc differences 
#emm = emmeans(model, ~ Treatment*Colony*DateMeasured)
#cld(emm, Letters=c(LETTERS)) #letter display
#pairs(emm)


```


## Count observations 

Calculate the number of measurements grouped by date, treatment and colony
```{r}
pam_data_spat %>%
  group_by(date, colony, condition) %>%
  summarise(observations = n())

pam_data_spat %>%
  group_by(date, colony, condition) %>%
  summarise(observations = n())
```





