---
title: "PAM data analysis"
author: "Jill Ashey"
date: "2023-08-14"
output: html_document
---

This code analyzes the pulse amplitude fluometry (PAM) data that was collected from P. asteriodes larvae and spat in Bermuda 2023. 

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(car)
library(emmeans)
library(multcomp)
library(knitr)
library(kableExtra)
library(lme4)
library(GGally)
library(lmerTest)
library(cowplot)
library(Hmisc)
library(ggstatsplot)
```

## First, I am going to analyze the data from the larval samples. 

## Load data 
```{r}
pam_data <- read.csv("data/pam/pam_larvae_dark.csv", na.strings=c("","NA"))
head(pam_data)
tail(pam_data)
```

I updated the data so that the pam larvae csv included DateCollected, DateMeasured, TimeMeasured, Colony, Basin, and Squarical. Now we can do plotting and stats on this data!

## Manipulate data frame 

Remove space from condition and colony columns
```{r}
pam_data$Colony <- sub(" ", "", pam_data$Colony)
```

Remove NAs from colony 
```{r}
pam_data <- pam_data %>%
  drop_na(Colony)
```

Set date
```{r}
pam_data$DateCollected <- as.POSIXct(pam_data$DateCollected, format = "%m/%d/%y")
pam_data$DateMeasured <- as.POSIXct(pam_data$DateMeasured, format = "%m/%d/%y")
```

## Make plots, identify outliers and do statistics 

### Fv.Fm

Plot Fv/Fm by treatment, colony and date
```{r}
ggplot(pam_data, aes(x = Colony, y = Fv.Fm, color = Treatment)) +
  geom_boxplot(aes(colour=Treatment,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw() +
  facet_wrap(~DateMeasured)
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

Some of the data only have one point...We may not be able to use these measurements if the measurement is n=1. 

Do stats on Fv/Fm by treatment, colony and date 
```{r}
model <- aov(Fv.Fm ~ Treatment*Colony*DateMeasured, data = pam_data)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ Treatment*Colony*DateMeasured, data = pam_data)
# anova assumptions met 

# View post-hoc differences 
emm = emmeans(model, ~ Treatment*Colony*DateMeasured)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)

# confusion
```


Plot Fv/Fm by treatment and colony
```{r}
ggplot(pam_data, aes(x = Colony, y = Fv.Fm)) +
  geom_boxplot(aes(colour=Treatment), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw()
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

Some of the data only have one point...We may not be able to use these measurements if the measurement is n=1. 

Do stats on Fv/Fm by treatment and colony 
```{r}
model <- aov(Fv.Fm ~ Treatment*Colony, data = pam_data)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ Treatment*Colony*DateMeasured, data = pam_data)
# anova assumptions met 

# View post-hoc differences 
TukeyHSD(model)

emm = emmeans(model, ~ Treatment*Colony)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)
```

Plot by treatment only
```{r}
ggplot(pam_data, aes(x = Treatment, y = Fv.Fm)) +
  geom_boxplot(aes(colour=Treatment,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw() 
  #facet_wrap(~date)
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

Do stats on Fv/Fm by treatment 
```{r}
model <- aov(Fv.Fm ~ Treatment, data = pam_data)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ Treatment*Colony*DateMeasured, data = pam_data)
# anova assumptions met 

# View post-hoc differences 
TukeyHSD(model)

emm = emmeans(model, ~ Treatment*Colony)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)
```

## Count observations 

Calculate the number of measurements grouped by date, treatment and colony
```{r}
summary_data <- pam_data %>%
  group_by(DateMeasured, Colony, Treatment) %>%
  summarise(observations = n())
```

We may only be able to use the measurements with >3 data points per date, treatment and colony if that makes sense. 














 


## Now, I am going to analyze the data from the spat samples. 

## Load data 
```{r}
pam_data <- read.csv("data/pam/pam_spat_measurements.csv", na.strings=c("","NA"))
head(pam_data)
tail(pam_data)
```

First look at the data - don't have info about their 'chamber' or tank id. 

## Manipulate data frame 

Remove space from condition and colony columns
```{r}
pam_data$condition <- sub(" ", "", pam_data$condition)
pam_data$colony <- sub(" ", "", pam_data$colony)
```

Remove NAs from tank and colony 
```{r}
pam_data <- pam_data %>%
  drop_na(condition, colony)
```
No data points removed here. 

Set date
```{r}
pam_data$date <- as.POSIXct(pam_data$date, format = "%m/%d/%y")
```

## Make plots, identify outliers and do statistics 

### Fv.Fm

Plot Fv/Fm
```{r}
ggplot(pam_data, aes(x = colony, y = Fv.Fm, color = condition)) +
  geom_boxplot(aes(colour=condition,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw() +
  facet_wrap(~date)
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

NEED TO LOOK AT OUTLIERS STILL 

Not sure how to do the ANOVA for this... maybe 3 way anova?? 
Do stats - run a two-way ANOVA
```{r}
model <- aov(Fv.Fm ~ condition*colony*date, data = pam_data)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ condition*colony*date, data = pam_data)
```

There is a violation of variance assumptions. Conduct non-parametric test 
```{r}
kruskal.test(Fv.Fm ~ colony, data = pam_data)
kruskal.test(Fv.Fm ~ condition, data = pam_data)
kruskal.test(Fv.Fm ~ date, data = pam_data)

# View post-hoc differences 
emm = emmeans(model, ~ condition*colony*date)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)
```

What if I plotted Fv.Fm for spat without the date? 
```{r}
ggplot(pam_data, aes(x = colony, y = Fv.Fm, color = condition)) +
  geom_boxplot(aes(colour=condition,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw() 
  #facet_wrap(~date)
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()

ggplot(pam_data, aes(x = condition, y = Fv.Fm)) +
  geom_boxplot(aes(colour=condition,), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw() 
  #facet_wrap(~date)
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

```{r}
model <- aov(Fv.Fm ~ condition*colony, data = pam_data)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ condition*colony*date, data = pam_data)
```

```{r}
t.test(Fv.Fm ~ condition, data = pam_data)
```







