---
title: "PAM data analysis"
author: "Jill Ashey"
date: "2023-08-14"
output: html_document
---

This code analyzes the pulse amplitude fluometry (PAM) data that was collected from P. asteriodes larvae and spat in Bermuda 2023. 

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(car)
library(emmeans)
library(multcomp)
library(knitr)
library(kableExtra)
library(lme4)
library(GGally)
library(lmerTest)
library(cowplot)
library(Hmisc)
library(ggstatsplot)
```

First, I am going to analyze the data from the larval samples. 

## Load data 
```{r}
pam_data <- read.csv("data/pam/jenny_larvae_measurements.csv", na.strings=c("","NA"))
head(pam_data)
tail(pam_data)
```

First look at the data - I'm not sure what the number refers to. It's labeled as Jenny 1-8, but does this correspond to colony? Not sure. Additionally, not all colonies are identified (some are NAs) and not all tanks are identified. I will remove these values for now, as I have no way of knowing what tank or colony the larvae came from. I will go back through the notebook pictures to see if I can figure it out. 

## Manipulate data frame 

Add treatment as a column
```{r}
pam_data <- pam_data %>%
  mutate(Treatment = case_when(
    tank == "1" ~ "Ambient",
    tank == "2" ~ "High", 
    tank == "3" ~ "High", 
    tank == "4" ~ "Ambient"
  ))
```

Remove NAs from tank and colony 
```{r}
pam_data <- pam_data %>%
  drop_na(tank, colony)
```
This got rid of 91 measurements, so we definitely need that data to make a decision about moving forward. Dropping of the NAs also got rid of the other dates, so we are now only working with 7/18/23. 

## Make plots, identify outliers and do statistics 

### Fv.Fm

Plot Fv/Fm
```{r}
ggplot(pam_data, aes(x = colony, y = Fv.Fm)) +
  geom_boxplot(aes(colour=Treatment), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw()
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

Identify outliers 
```{r}
outlier.plot <- ggbetweenstats(pam_data, colony, Fv.Fm, outlier.tagging = T)
outlier.plot

# Don't seem to be any major outliers 
```

Do stats - run a two-way ANOVA
```{r}
model <- aov(Fv.Fm ~ Treatment*colony, data = pam_data)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ Treatment*colony, data = pam_data)
```

No significant differences 

### Fo

Plot Fo
```{r}
ggplot(pam_data, aes(x = colony, y = Fo)) +
  geom_boxplot(aes(colour=Treatment), width=.5, outlier.shape= NA) +
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red"))+
  theme_bw()
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

Identify outliers 
```{r}
outlier.plot <- ggbetweenstats(pam_data, Treatment, Fo, outlier.tagging = T)
outlier.plot

### NEED TO REMOVE OUTLIERS 
```

Do stats - run a two-way ANOVA
```{r}
model <- aov(Fo ~ Treatment*colony, data = pam_data)
summary(model)
qqPlot(residuals(model))
leveneTest(residuals(model) ~ Treatment*colony, data = pam_data)
```
There is a violation of variance assumptions. Conduct non-parametric test 
```{r}
kruskal.test(Fo ~ colony, data = pam_data)
kruskal.test(Fo ~ Treatment, data = pam_data)

## Need to figure out how to do this in a two-way fashion 
```




