--
title: "Larval Photosynthesis Rates Extraction with LoLinR - Bermuda 2023 - PI Curves"
author: "Jill Ashey"
date: "2023-07-13"
output: html_document
---

This script is extracting photosynthesis rates under different light levels for P. asteroides larvae in Bermuda. The code is based on A. Huffmyer's PI [script](https://github.com/AHuffmyer/larval_symbiont_TPC/blob/main/scripts/larval_pi_curve_sdr_extraction.Rmd). Important note - this data was collected in 4 mL vials, not the 24-well plates.

## Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
library(devtools)
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr') 
if ("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc') 
if ("respR" %in% rownames(installed.packages()) == 'FALSE') install.packages('respR') 


#load packages
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('plyr')
library('dplyr')
library('stringr')
library('Rmisc')
library('respR')
library("tidyverse")
```

## Read in files  

Set the path of all respirometry files within the R project.   

```{r, warning=FALSE, message=FALSE}
path.p<-"../data/larval_pi_curves/runs" #location of files
```

Bring in the file names.
```{r, warning=FALSE, message=FALSE}
# bring in the respiration file names
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) 
# remove test file once we have more pi curve data 
```

Bring in the run metadata. 
```{r}
run_metadata<-read.csv("../data/larval_pi_curves/larval_PIcurve_SDR_Run_Info.csv")
```

## Extract photosynthesis rates  

Generate photosynthesis data frames.

```{r, warning=FALSE, message=FALSE}
#generate a 7 column dataframe with specific column names
#photosynthesis
Photo.R <- data.frame(matrix(NA, ncol=7))
colnames(Photo.R) <- c("Date", "Plate","Sample.ID","Chamber.ID","Light", "Intercept", "umol.L.min")

Photo.Rb <- data.frame(matrix(NA, ncol=7))
colnames(Photo.Rb) <- c("Date", "Plate","Sample.ID","Chamber.ID","Light", "Intercept", "umol.L.min")
```

Load in the sample information file. It is important to have chambers in the order "A1, B1, C1, D1", rather than "A1, A2, A3, A4".  

Read in sample and run information and merge into a master metadata file. 
```{r}
# Load PI curve sample metadata (i.e., which larvae were in which runs)
Sample.Info <- read_csv(file = "../data/larval_pi_curves/larval_PIcurve_SDR_Sample_Info.csv")%>%
  select(!TimeStart)%>%
  select(!TimeEnd)%>%
  select(!Notes)

rename <- Sample.Info$Chamber.ID
rename_trunc <- rename[9:12]
samp <- Sample.Info$Sample.ID

plate <- as.numeric(str_sub(file.names, 28, 29))
date <- str_sub(file.names, 1, str_length(file.names)-28) #grab date from file name

# Load PI curve run metadata (i.e., light levels and interval times for each run)
Run.Info <- read_csv(file = "../data/larval_pi_curves/larval_PIcurve_SDR_Run_Info.csv")

Light_Values <- Run.Info$Light_Level

# Join all coral and run metadata
metadata <- full_join(Sample.Info, Run.Info) %>%
  mutate(Date = as_date(as.character(Date), format = "%Y%m%d"))

starttime<-Run.Info$IntervalStart
endtime<-Run.Info$IntervalStop
```

Script is having an issue extracting the plate number from the file name. 

Remove "Type" columns that have NA in them 
```{r}
metadata <- metadata %>%
  drop_na(Type)
```


Run loop to extract slopes from photosynthesis data from each light level in each file name/plate. 

```{r, results=FALSE, warning=FALSE, message=FALSE}

for(file in 1:length(file.names)) { # for every file in list start at the first and run this following function
  
  for (i in Light_Values) { #in every file, for each light value 
    
  Photo.Data <-read.table(file.path(path.p,file.names[file]), skip = 56, header=T, sep=",", na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Photo.Data$Temp <- Photo.Data[,31] #assigns temp column
  #Photo.Data$Time.Min. <- seq.int(from=0, to=((nrow(Photo.Data)*0.25)-0.25), by = 0.25) #set time in min
  
  Photo.Data <- Photo.Data[,-c(3:10,15:26)] #remove unused columns
  
  # extract start and end times for the respective plate 
  
  starttime<-Run.Info%>%
    select(Plate, Light_Level, IntervalStart)%>%
    filter(Plate==plate[file])%>%
    filter(Light_Level==i)%>%
    select(IntervalStart)%>%
    as.data.frame()

   starttime<-starttime[1,1]
   
   endtime<-Run.Info%>%
    select(Plate, Light_Level, IntervalStop)%>%
    filter(Plate==plate[file])%>%
    filter(Light_Level==i)%>%
    select(IntervalStop)%>%
    as.data.frame()

   endtime<-endtime[1,1]
   
  #filter by light interval - need to bind in dataframe to tell it which start and end time to use 
  Photo.Data <- Photo.Data %>% #filters data by interval for light 
     filter(Time.Min. > starttime)%>%
     filter(Time.Min. < endtime) 
  
  Photo.Data.N <- Photo.Data[,3:6] #subset desired columns
  
  #add column names back in 
  Photo.Data.N<-as.data.frame(Photo.Data.N)
  
  for(j in 1:(ncol(Photo.Data.N))){
    model <- rankLocReg(
      xall=Photo.Data$Time.Min., yall=as.numeric(Photo.Data.N[, j]), 
      alpha=0.4, method="pc", verbose=TRUE) #extract slopes, percentile rank method with minimum window size of 0.4. This means that in order to fit a slope, it has to encompass at least 40% of available datapoints. 
    
    pdf(paste0("../output/larval_pi_curves/PhotosynthesisPlots/",date[file], "_Plate",plate[file],"_",rename_trunc[j],"_light", Light_Values[i],"_regression_trunc.pdf")) #generate output file names
    plot(model)
    dev.off()
    
    Photo.Rb[j,1] <- as.character(date[file]) #stores the date
    Photo.Rb[j,2] <- as.character(plate[file]) #stores the run number
    Photo.Rb[j,3] <- as.character(samp[j+(i-1)*ncol(Photo.Data.N)]) #stores the sample ID
    Photo.Rb[j,4] <- as.character(rename_trunc[j]) #stores the chamber ID
    Photo.Rb[j,5] <- as.character(i) #stores the chamber ID
    Photo.Rb[j,6:7] <- model$allRegs[i,c(4,5)] #inserts slope and intercept in the dataframe
    
  }
  
  Photo.R <- rbind(Photo.R, Photo.Rb) #bind final data frame
  }
  
}
```

JA: There is some issue with the start/end times and the Time.Min. I think the issue is somewhere between lines 134-157, but I'm not totally sure. Every time I try to run the chunk above, it gives me this error: "Error in combn(lenX, 2) : n < m", which makes me think that the issue is with the time filtering step. Again, not totally sure. Might be best to talk to AH about this. 

ZD: Got it working by removing the NA columns.  The issue was that we deleted certain rows where the data was incorrectly recorded, so trying to recode the Time.Min. column using the sequence code "#Photo.Data$Time.Min. <- seq.int(from=0, to=((nrow(Photo.Data)*0.25)-0.25), by = 0.25) #set time in min" was shifting the times by incorrect values. I have commented out these lines here and below and will attempt to run the script this way.

Calculate average temperature of each run and export to a table to confirm temperatures are not different between PAR values.    
```{r}
# list files 
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) 

#generate matrix to populate
Temp.P <- data.frame(matrix(NA, ncol=4))
colnames(Temp.P) <- c("Date", "Plate","Temp.C", "Light")

Temp.Pb <- data.frame(matrix(NA, ncol=4))
colnames(Temp.Pb) <- c("Date", "Plate","Temp.C", "Light")

#read in temps and generate mean values for each step of the light curve  
for(file in 1:length(file.names)) { 
  
  for(i in Light_Values) {# for every file in list start at the first and run this following function
  
  Temp.Data <-read.table(file.path(path.p,file.names[file]), skip = 56, header=T, sep=",", na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Temp.Data$Temp <- Temp.Data[,31] #assigns temp column
  #Temp.Data$Time.Min. <- seq.int(from=0, to=((nrow(Temp.Data)*0.25)-0.25), by = 0.25) #set time in min
  
  Temp.Data <- Temp.Data[,-c(3:10,15:26)] #remove unused columns
  
  
  starttime<-Run.Info%>%
    select(Plate, Light_Level, IntervalStart)%>%
    filter(Plate==plate[file])%>%
    filter(Light_Level==i)%>%
    select(IntervalStart)%>%
    as.data.frame()

   starttime<-starttime[1,1]
   
   endtime<-Run.Info%>%
    select(Plate, Light_Level, IntervalStop)%>%
    filter(Plate==plate[file])%>%
    filter(Light_Level==i)%>%
    select(IntervalStop)%>%
    as.data.frame()

   endtime<-endtime[1,1]
   
  #filter by light interval - need to bind in dataframe to tell it which start and end time to use 
  Temp.Data <- Temp.Data %>% #filters data by interval for light 
     filter(Time.Min. > starttime)%>%
     filter(Time.Min. < endtime) 
  
    Temp.Pb[j,1] <- as.character(date[file]) #stores the date
    Temp.Pb[j,2] <- as.character(plate[file]) #stores the run number
    Temp.Pb[j,3] <- mean(Temp.Data$Temp) #stores the sample ID
    Temp.Pb[j,4] <- mean(Light_Values[i]) #stores the sample ID
    
    
    Temp.P <- rbind(Temp.P, Temp.Pb) #bind final data frame
    Temp.P <- na.omit(Temp.P)
  
}

}
write.csv(Temp.P, paste0("../output/larval_pi_curves/runs_temp.csv")) #save respiration rate data
```


Save photosynthesis data frames.  

```{r, results=FALSE, warning=FALSE, message=FALSE}
Photo.R <- Photo.R[-1,] #remove empty row

write.csv(Photo.R, paste0("../output/larval_pi_curves/larval_pi_curves_sdr_rates_raw.csv")) #save respiration rate data

plot(Photo.R$umol.L.min~as.factor(Photo.R$Plate)*as.factor(Photo.R$Light), side = 2, las = 2, xlab="" )

```

```{r, warning=FALSE, message=FALSE}
Photo.Rates <- read.csv(file = "../output/larval_pi_curves/larval_pi_curves_sdr_rates_raw.csv") #read file back in so slopes don't have to be generated every time
Photo.Rates = subset(Photo.Rates, select = -c(X) ) #remove empty column

#format "run" column
Photo.Rates<-Photo.Rates %>%
  mutate(Plate = str_sub(Plate, 1, -1))
Photo.Rates$Plate<-as.integer(Photo.Rates$Plate) #format as # rather than run #, set as integer
```

## Standardize and normalize  

Merge rates files with sample info for testing and manipulation.  

```{r, warning=FALSE, message=FALSE}

oxygen<-full_join(Sample.Info, Photo.Rates) #add photosynthesis data
colnames(oxygen)[colnames(oxygen) == 'Intercept'] <- 'Photo.Intercept' #rename to specify R
colnames(oxygen)[colnames(oxygen) == 'umol.L.min'] <- 'P.umol.L.min' #rename to specify R

#make unique ID column
oxygen$unique<-paste(oxygen$Plate,"-",oxygen$Chamber.ID)
Sample.Info$unique<-paste(Sample.Info$Plate,"-",Sample.Info$Chamber.ID)

oxygen$Sample.ID<-Sample.Info$Sample.ID[match(oxygen$unique, Sample.Info$unique)]
oxygen$SDR<-Sample.Info$SDR[match(oxygen$unique, Sample.Info$unique)]
oxygen$Tank<-Sample.Info$Tank[match(oxygen$unique, Sample.Info$unique)]
oxygen$Symbiont<-Sample.Info$Symbiont[match(oxygen$unique, Sample.Info$unique)]
oxygen$Type<-Sample.Info$Type[match(oxygen$unique, Sample.Info$unique)]
oxygen$Salinity<-Sample.Info$Salinity[match(oxygen$unique, Sample.Info$unique)]
oxygen$Run<-Sample.Info$Run[match(oxygen$unique, Sample.Info$unique)]
oxygen$Volume<-Sample.Info$Volume[match(oxygen$unique, Sample.Info$unique)]
oxygen$Org.Number<-Sample.Info$Org.Number[match(oxygen$unique, Sample.Info$unique)]
```

Remove samples that have inaccurate slope extraction for respiration rates (determined by PDF output files)  


Account for volume to obtain umol per minute.   

```{r, results=TRUE, warning=FALSE, message=FALSE}
#Account for chamber volume to convert from umol L-1 m-1 to umol m-1. This removes per Liter
oxygen$P.umol.min <- oxygen$P.umol.L.min * oxygen$Volume #calculate

plot(as.factor(oxygen$Plate), oxygen$P.umol.L.min, xlab="Plate", ylab="rate") #photo rates
```

Subtract blank values. Average blank calculated for each run. Display mean blank value.  
```{r}
blank_data <- subset(oxygen, Type == "Blank") #subset to blank data only
plot(as.factor(blank_data$Plate), blank_data$P.umol.min, xlab="Plate", ylab="umol O2 min-1") #blanks during dark phase

plot(as.factor(blank_data$Light), blank_data$P.umol.min, xlab="Plate", ylab="umol O2 min-1") #blanks during dark phase

#filter blank values 
#blank_data <- blank_data%>%
#  filter(P.umol.min> -0.0004)%>%
#  filter(P.umol.min < 0.0003)

plot(as.factor(blank_data$Plate), blank_data$P.umol.min, xlab="Plate", ylab="umol O2 min-1") #blanks during dark phase

#display mean blank values
mean(blank_data$P.umol.min, na.rm=TRUE) #mean P phase blanks

photo.blnk <- aggregate(P.umol.min ~ Date*Plate, data=blank_data, mean) #calculate average blank during light for each run
colnames(photo.blnk)[colnames(photo.blnk) == 'P.umol.min'] <- 'P.Blank.umol.min' #rename to specify blank for R

oxygen <- full_join(oxygen, photo.blnk) #add R blanks to master
```

ZD: I had to remove the step below because our blank values did not fall within this range. Want to confirm this with Jill / Ariana.

"
  #filter blank values 
  #blank_data <- blank_data%>%
  #  filter(P.umol.min> -0.0004)%>%
  #  filter(P.umol.min < 0.0003)
"

Subtract blank values to generate a "corrected" value for umol O2 min-1.  

```{r, warning=FALSE, message=FALSE}
oxygen$P.umol.min.corr<-oxygen$P.umol.min-oxygen$P.Blank.umol.min #subtract R blanks
```

```{r}
plot(oxygen$P.umol.min.corr)
```

Normalize to biologically relevant measure. Here, normalize to number of larvae. This can be substituted or changed for larval size/volume as well.  

```{r, warning=FALSE, message=FALSE}
oxygen.bio <- oxygen %>% filter(Type == "Sample") #isolate only biological samples and drop unused factor levels 
oxygen.bio <- droplevels(oxygen.bio) #drop unused factor levels

#respiration
oxygen.bio$P.umol.org.min <- oxygen.bio$P.umol.min.corr/oxygen.bio$Org.Number #calculate oxygen per organism 
oxygen.bio$P.nmol.org.min <- oxygen.bio$P.umol.org.min*1000 #calculate nanomoles 

```

Plot values.  
```{r}
plot(oxygen.bio$P.nmol.org.min)
```

Add in relevant metadata. 
```{r}
oxygen.bio$PAR <- Run.Info$Light_Value[match(oxygen.bio$Light, Run.Info$Light_Level)]
```

Save as .csv file.  
```{r, warning=FALSE, message=FALSE}
write.csv(oxygen.bio, paste0("../output/larval_pi_curves/calculated_normalized_photo_rates.csv")) #save final file
```


